---
title: "Sets of SNPs"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
editor_options: 
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5,fig.height = 4,cache=FALSE)
options(width = 300)
library(knitr)
library(tidyverse)
library(kableExtra)
```

In this document, we create a list of three sets of SNPs that will be used to calculate the genomic offset.

  - The **reference SNPs**.

  - the **common candidates**: with *both* BF > 5dB in BayPass and SNP loadings > 3 sd cutoff in the RDA for the chosen set of covariates.
  
  - the **candidates under expected strong selection**: RDA candidates that show a strong association with at least one covariate, i.e. with $\beta_{RDA} > 0.3$, and all the BayPass candidates (BF > 5dB).

  - the **merged candidates**: with *either* BF > 5dB in BayPass or SNP loadings > 3 sd cutoff in the RDA for the chosen set of covariates.
                

# Load data

Load the entire genomic dataset (for the reference SNPs).

```{r LoadGenomicData}
geno <- readRDS(file="data/GenomicData/FormattedData/GenomicData_454clones_9817snps.rds")
```


> Load the Baypass candidates.

```{r LoadCandBayPass}
modelBayPass <- "IS" # "IS" or "MCMC"
candBaypass <- readRDS(file=paste0("outputs/CandidateSNPs/baypass/CandSNPsBayPass",modelBayPass,".rds"))
candBaypass  %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("hover", "condensed"), full_width = F)
```

Column meaning:

  - `medianBF`: median estimate of the Bayes factor from BayPass.
  
  - `medianBeta`: median estimate of the $\beta$ estimate from BayPass.
  
  - `COVARIABLE`: covariate with which the SNP was found to be associated with.

> Load the RDA candidates.

```{r LoadCandRDA}
candRDA <- readRDS(file="outputs/CandidateSNPs/rda/352CandSNPsRDA.rds")
candRDA[1:10,]  %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("hover", "condensed"), full_width = F)
```

Column meaning:

  - `axis`: Significant axis. (comment: a snp can have several significant axes but, as we removed the duplicate SNPs, only one is noted here)
  
  - `loading`: SNP loading on the significant axis.
  
  - `bio5` ...  `BurnedArea`: correlation with each environmental covariate.
  
  - `predictor`: main predictor for each SNP.
  
  -  `correlation`: correlation with the main predictor.

# Merged candidates



```{r MergedCand}
mergedcand <- candBaypass %>% 
  filter(COVARIABLE %in% c("bio5","bio6","bio12","bio15","water_top","depth_roots","TRI","BurnedArea")) %>% 
  full_join(candRDA,by="snp")  %>% 
  select(snp,COVARIABLE,medianBF,medianBeta,loading,contains("bio"),water_top,depth_roots,BurnedArea,TRI) %>% 
  dplyr::rename(varBaypass=COVARIABLE,
         BF.Baypass=medianBF,
         beta.Baypass=medianBeta,
         loading.RDA=loading,
         beta.bio5.RDA=bio5,
         beta.bio6.RDA=bio6,
         beta.bio12.RDA=bio12,
         beta.bio15.RDA=bio15,
         beta.water_top.RDA=water_top,
         beta.depth_roots.RDA=depth_roots,
         beta.BurnedArea.RDA=BurnedArea,
         beta.TRI.RDA=TRI) 

mergedcand %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("hover", "condensed"), full_width = F)
```

There are **`r length(mergedcand$snp)` merged candidates.**

# Candidates under expected strong selection

```{r CandSNPsUnderStrongSelection}
highcand <- candBaypass %>% 
  filter(COVARIABLE %in% c("bio5","bio6","bio12","bio15","water_top","depth_roots","TRI","BurnedArea")) %>% 
  full_join(candRDA,by="snp") %>% 
  filter(abs(medianBeta) >0 |abs(bio5)>0.3 | abs(bio6)>0.3 | abs(bio12)>0.3 | abs(bio15)>0.3 | abs(water_top)>0.3 | abs(depth_roots)>0.3 | abs(BurnedArea)>0.3 | abs(TRI)>0.3) %>% 
  select(snp,COVARIABLE,medianBF,medianBeta,loading,contains("bio"),water_top,depth_roots,BurnedArea,TRI) %>% 
  dplyr::rename(varBaypass=COVARIABLE,
         BF.Baypass=medianBF,
         beta.Baypass=medianBeta,
         loading.RDA=loading,
         beta.bio5.RDA=bio5,
         beta.bio6.RDA=bio6,
         beta.bio12.RDA=bio12,
         beta.bio15.RDA=bio15,
         beta.water_top.RDA=water_top,
         beta.depth_roots.RDA=depth_roots,
         beta.BurnedArea.RDA=BurnedArea,
         beta.TRI.RDA=TRI) 

highcand %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("hover", "condensed"), full_width = F)
```


There are **`r length(highcand$snp)` candidates under expected strong selection**.

# Common candidates


```{r ComCand}
comcand <- candBaypass %>% 
  filter(COVARIABLE %in% c("bio5","bio6","bio12","bio15","water_top","depth_roots","TRI","BurnedArea")) %>% 
  inner_join(candRDA,by="snp") %>% 
  select(snp,COVARIABLE,medianBF,medianBeta,loading,contains("bio"),water_top,depth_roots,BurnedArea,TRI) %>% 
  dplyr::rename(varBaypass=COVARIABLE,
         BF.Baypass=medianBF,
         beta.Baypass=medianBeta,
         loading.RDA=loading,
         beta.bio5.RDA=bio5,
         beta.bio6.RDA=bio6,
         beta.bio12.RDA=bio12,
         beta.bio15.RDA=bio15,
         beta.water_top.RDA=water_top,
         beta.depth_roots.RDA=depth_roots,
         beta.BurnedArea.RDA=BurnedArea,
         beta.TRI.RDA=TRI) 

comcand %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("hover", "condensed"), full_width = F)

```

There are **`r length(comcand$snp)` candidates under expected strong selection**.

# Exporting candidate SNPs

## List of sets of candidate SNPs

```{r CreatingListSnpSets}
snp.sets <- list(rownames(geno), # Reference SNPs
                 mergedcand$snp, # Merged candidates
                 highcand$snp,   # Candidates under expected strong selection
                 comcand$snp)    # Common candidates

names(snp.sets) <- c("Ref","Mer","Mid","Com")
#saveRDS(snp.sets,file=paste0("outputs/ListSetsSNPs",modelBayPass,".rds"))
```

## DF with candidate SNP codes

Below we create a dataframe with all the candidate SNPs and their associated SNP codes (original ID, ID in the  Axiom assay and ID in the Illumina Infinium assay).

<!-- DF sent to Katha the 2sn of June 2022. -->

```{r DFCandSNPwithSNPcodes}
snp.codes <- readRDS("data/GenomicData/FormattedData/SNPcodes.rds") %>% 
  dplyr::rename(snp_ID_Juliette=snp_ID)

list.snps.cand <- list(mergedcand, highcand, comcand)
names(list.snps.cand) <- c("Mer","Mid","Com")
df.snps <- list.snps.cand %>% 
  bind_rows(.id="SnpSet") %>% 
  dplyr::select(-loading.RDA) %>% 
  dplyr::rename(snp_ID_Juliette=snp) %>% 
  left_join(snp.codes,by="snp_ID_Juliette")
  
saveRDS(df.snps,file=paste0("outputs/ListSetsCandSNPs_WithSNPCodes_",modelBayPass,".rds"))
write.csv(df.snps, file= paste0("outputs/ListSetsCandSNPs_WithSNPCodes_",modelBayPass,".csv"), row.names = F)
```

Column meaning:

  - `SnpSet`: set of candidate SNPs in which the SNP is included (Mer for merged candidates, Mid for the candidates under expected strong selection and Com for the common candidates)
  
  - `snp_ID_Juliette`: My own snp code I used in my analyses.

  - `varBaypass`: the covariate with which the SNP was found to be associated with.

  - `BF.Baypass`: median estimate of the Bayes factor in the BayPass analysis.
  
  - `beta.Baypass`: median estimate of the beta coefficient (standing for the linear association between the SNP frequency and the environmental covariate) in the BayPass analysis.

  - `beta.(covariate).RDA`: coefficient of correlation between the SNP and each environmental covariate in the RDA.
  
  - `original_ID`: original SNP ID.

  - `affx_ID`:  ID from the Axiom assay.

  - `infinium_ID`: ID from the Illumina Infinium assay.

`r knitr::opts_chunk$set(eval = F)`
