---
title: "Vizualisation of environmental data and building rasters"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: kable
    highlight: textmate
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
    toc_depth: '4'
editor_options:
  chunk_output_type: console
always_allow_html: true
---


<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=F)
options(width = 300)
library(knitr)      # CRAN v1.26
library(kableExtra) # CRAN v1.1.0
library(dplyr)      # CRAN v1.0.0
library(psych)      # CRAN v1.8.12 (pairs.panels function)
library(tidyverse)  # CRAN v1.3.0
library(raster)     # CRAN v3.3-13
library(stringr)    # CRAN v1.4.0
library(ggbiplot)
```



# Vizualisation of environmnent data

Here the set of environmental covariates we use in the present study:

Burned area: `BurnedArea`, i.e. the average of the monthly burned area from June 1995 to December 2014 (hectares).

Topographic variable:`TRI`: topographic ruggedness index (unitless)

Soil-related variables from the European Soil Database Derived data:

  - `water_top` = the total available water content (mm)
  - `depth_roots` = depth available to roots (cm)

Climatic variables from WordClim (average for the years 1970-2000):

  - `bio5`: max temperature of the warmest month (°C)
  - `bio6`: min temperature of the coldest month (°C)
  - `bio12`: annual precipitation (mm)
  - `bio15`: precipitation Seasonality (coefficient of Variation)


This set of covariates was selected because:

  - in previous analyses based on a subset of the genomic data (Infinium assay data), this is the set (among 6 sets tested) that shows the highest number of candidate SNPs. 
  
  - this set has biological sense. In particular, it includes `bio6`, the minimum temperature of the coldest month, and studies are accumulating showing that maritime pine patterns of adaptation to climate are associated with minimum temperatures. 
  
```{r CorrelationsPanels,fig.height=9,fig.width=12}
# We load the dataset with the extracted environmental variables:
data <- read.csv("data_DRYAD/PhenoEnvPopStructureDataset.csv")

select.var <- c("bio5WC_prov",
                "bio6WC_prov",
                "bio12WC_prov",
                "bio15WC_prov",
                "water_top_prov",
                "depth_roots_prov",
                "TRI",
                "BurnedArea")

data %>% 
  dplyr::select(all_of(select.var)) %>% 
  unique() %>% 
  dplyr::rename(bio5="bio5WC_prov",
                bio6="bio6WC_prov",
                bio12="bio12WC_prov",
                bio15="bio15WC_prov",
                water_top="water_top_prov",
                depth_roots="depth_roots_prov") %>% 
  pairs.panels(scale=T,hist.col="palegreen1")
```

Other visualizations:

```{r PlotCorr,fig.height=7.3,fig.width=7.3}
source("scripts/Functions/corpmat.R")
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))


PlotCors <- function(data,select.var,nbsize){
  cor <- cor(data[,select.var])

# matrix of the p-value of the correlation
p.mat <- corpmat(cor)

corrplot::corrplot(cor, method="color", col=col(200),  
                   type="upper", order="hclust", 
                   addCoef.col = "black", # Add coefficient of correlation
                   tl.col="black", tl.srt=45, #Text label color and rotation
                   # Combine with significance
                   p.mat = p.mat, sig.level = 0.01, insig = "blank", 
                   # hide correlation coefficient on the principal diagonal
                   diag=FALSE,number.cex=nbsize)

pca <- prcomp(data[,select.var], center = TRUE,scale. = TRUE)

ggbiplot(pca,varname.size =4) +  ylim(-2.5, 2.5) +    xlim(-3, 3) +  
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(size=12))
}

data %>% 
  dplyr::select(all_of(select.var)) %>% 
  unique() %>% 
  dplyr::rename(bio5="bio5WC_prov",
                bio6="bio6WC_prov",
                bio12="bio12WC_prov",
                bio15="bio15WC_prov",
                water_top="water_top_prov",
                depth_roots="depth_roots_prov") %>% 
  PlotCors(select.var = colnames(.),nbsize = 0.6)
```

# Creating raster stacks of environmental variables


Climatic variables:

```{r ClimaticVariables}
clim.var <- c("bio5","bio6","bio12","bio15")
```

```{r FunctionRemoveNAsStack}
remove.NAs.stack<-function(rast.stack){
  nom<-names(rast.stack)
  test1<-calc(rast.stack, fun=sum)
  test1[!is.na(test1)]<-1
  test2<-rast.stack*test1
  test2<-stack(test2)
  names(test2)<-nom
  return(test2)
}
```


Load the mask of the maritime pine distribution (made with Euforgen data and the NFI plots)

```{r LoadMask}
PinpinDistri <- shapefile('data_DRYAD/Mapping/PinpinDistriEUforgen_NFIplotsBuffer10km.shp')
plot(PinpinDistri)
```


`r knitr::opts_chunk$set(eval = F)`

## Current climate (res 30 seconds)

We create raster stacks with the topographic, soil, fire-related variables and current climatic conditions from WorldClim (i.e. from 1970 to 2000).

```{r CurrentClimateRes30sec}

# Climatic rasters
# ----------------
grids <- list.files("data/climate/CurrentClimate/WorldClim30sec_1970_2000/")
grids <- grids[grepl("bio_5|bio_6|bio_12|bio_15",grids)==T]
grids <- c(grids[[3]],grids[[4]],grids[[1]],grids[[2]]) # reorder
rast.clim <- raster::stack(paste0("data/climate/CurrentClimate/WorldClim30sec_1970_2000/", grids))
names(rast.clim) <- clim.var
rast.clim <- crop(rast.clim,extent(PinpinDistri)) 
rast.clim <- mask(rast.clim,PinpinDistri, updatevalue=NA)


# Soil rasters
# ------------
rast.soil <- stack("data/soil/STU_EU_T_TAWC_WGS84.tif","data/soil/STU_EU_DEPTH_ROOTS_WGS84.tif")
names(rast.soil) <- c("water_top","depth_roots")
rast.soil <- raster::crop(rast.soil,extent(rast.clim))   # attribute the extent of climatic raster to soil raster
rast.soil <- raster::resample(rast.soil,rast.clim)       # same number of cells between the soil and climatic rasters
rast.soil <- mask(rast.soil,PinpinDistri,updatevalue=NA) # keep only cells within maritime pine distribution 


# TRI raster
# ----------
# > The models are fitted with TRI values from raster at 90m resolution, and the 
# projections will be on raster at 1km resolution (to have the same resolution as WorldClim variables).
rast.tri <- raster("data/Topography/TRI_WGS84.tif")
rast.tri <- crop(rast.tri,extent(rast.clim))           # attribute the extent of climatic raster to TRI rasters
rast.tri <- raster::resample(rast.tri,rast.clim)       # same number of cells between the TRI and climatic rasters
names(rast.tri) <- c("TRI")
rast.tri <- mask(rast.tri,PinpinDistri,updatevalue=NA) # keep only cells within maritime pine distribution 


# Burned Area raster
# ------------------
rast.BA <- raster("data/FireData/GFED4/PreparedRasters/GFED4_199507_201412_AllCountries.grd")
rast.BA <- crop(rast.BA,extent(rast.clim))
rast.BA <- raster::resample(rast.BA,rast.clim)
names(rast.BA) <- "BurnedArea"
rast.BA <- mask(rast.BA,PinpinDistri,updatevalue=NA)


# Merge and save the rasters in a stack
# ------------------------------------
stackall <- stack(rast.clim,rast.soil,rast.tri,rast.BA) # merge in a stack the different layers
stackall <- remove.NAs.stack(stackall) # same cells with NAs in all layers
stackall
writeRaster(stackall, filename="data/StacksEnvVars/Stack_CurrentClimate_Res30seconds.grd", format="raster",overwrite=TRUE)
```


## Future climate (res 2.5 minutes)

> Period: 2041-2060

Predictions of future climate come from the nine global climate models (GCMs) available in WorldClim for the period 2041 to 2060 and the Shared-socio Economic Pathways (SSP) 370 and 585 (https://www.worldclim.org/data/cmip6/cmip6climate.html).

For each global climate model, we create a **stack of 8 environmental variables**: the 4 variables describing the future climate (`bio5`, `bio6`, `bio12` and `bio15`), the soil-related variables (`water_top` and `depth_roots`), the topographic rugedness index (`TRI`) and the burned area (`BurnedArea`).

### SSP3-7.0 

```{r FutureClimateSSP370}
rast.to.crop <- raster("data/StacksEnvVars/Stack_CurrentClimate_Res30seconds.grd")

# The nine global climate models:
GCMs <- list.files("data/climate/FutureClimate/share/spatial03/worldclim/cmip6/7_fut/2.5m/")
GCMs

list.rast.GCMs <- sapply(GCMs, function(i){
  
  rast.clim <- stack(paste0("data/climate/FutureClimate/share/spatial03/worldclim/cmip6/7_fut/2.5m/",i,"/ssp370/wc2.1_2.5m_bioc_",i,"_ssp370_2041-2060.tif"))
  
  # I understand that these 19 variables here are the WorldClim bioclimatic variables averaged over the period 2041/2060.
  
  select.layers <- names(rast.clim)[grepl("2060.5$|2060.6|2060.12|2060.15",names(rast.clim))==T]
  rast.clim <- raster::subset(rast.clim,select.layers)
  names(rast.clim) <- clim.var
  #rast.clim <- rast.clim[[clim.var]]
  rast.clim <- raster::crop(rast.clim,extent(rast.to.crop))
  rast.clim <- mask(rast.clim,PinpinDistri, updatevalue=NA)
}, USE.NAMES = TRUE,simplify=FALSE)
  

# Soil rasters
rast.soil <- stack("data/soil/STU_EU_T_TAWC_WGS84.tif","data/soil/STU_EU_DEPTH_ROOTS_WGS84.tif")
names(rast.soil) <- c("water_top","depth_roots")
rast.soil <- raster::crop(rast.soil,extent(list.rast.GCMs[[1]]))
rast.soil <- raster::resample(rast.soil,list.rast.GCMs[[1]])
rast.soil <- mask(rast.soil,PinpinDistri,updatevalue=NA)
  
# TRI raster
rast.tri <- raster("data/Topography/TRI_WGS84.tif")
rast.tri <- crop(rast.tri,extent(list.rast.GCMs[[1]]))
rast.tri <- raster::resample(rast.tri,list.rast.GCMs[[1]])
names(rast.tri) <- c("TRI")
rast.tri <- mask(rast.tri,PinpinDistri,updatevalue=NA)
  
# Burned Area
rast.BA <- raster("data/FireData/GFED4/PreparedRasters/GFED4_199507_201412_AllCountries.grd")
rast.BA <- crop(rast.BA,extent(list.rast.GCMs[[1]]))
rast.BA <- raster::resample(rast.BA,list.rast.GCMs[[1]])
names(rast.BA) <- "BurnedArea"
rast.BA <- mask(rast.BA,PinpinDistri,updatevalue=NA)
  

for(i in GCMs){
  
  stackall <- stack(list.rast.GCMs[[i]],rast.soil,rast.tri,rast.BA)
  stackall <- remove.NAs.stack(stackall) # same cells with NAs in all layers
  
  writeRaster(stackall, 
              filename=paste0("data/StacksEnvVars/FutureClimate/2041to2060/SSP370/Stack_",str_remove_all(i,"-"),".grd"),
              format="raster",
              overwrite=TRUE)
  
}
```


### SSP5-8.5

```{r FutureClimateSSP585}
# We remove the GCM that is not included in this SSP:
GCMs <- GCMs[!GCMs=="GFDL-ESM4"]

list.rast.GCMs <- sapply(GCMs, function(i){
  
  rast.clim <- stack(paste0("data/climate/FutureClimate/share/spatial03/worldclim/cmip6/7_fut/2.5m/",i,"/ssp585/wc2.1_2.5m_bioc_",i,"_ssp585_2041-2060.tif"))
  
  # I understand that these 19 variables here are the WorldClim bioclimatic variables averaged over the period 2041/2060.
  
  select.layers <- names(rast.clim)[grepl("2060.5$|2060.6|2060.12|2060.15",names(rast.clim))==T]
  rast.clim <- raster::subset(rast.clim,select.layers)
  names(rast.clim) <- clim.var
  #rast.clim <- rast.clim[[clim.var]]
  rast.clim <- raster::crop(rast.clim,extent(rast.to.crop))
  rast.clim <- mask(rast.clim,PinpinDistri, updatevalue=NA)
}, USE.NAMES = TRUE,simplify=FALSE)
  

# Soil rasters
rast.soil <- stack("data/soil/STU_EU_T_TAWC_WGS84.tif","data/soil/STU_EU_DEPTH_ROOTS_WGS84.tif")
names(rast.soil) <- c("water_top","depth_roots")
rast.soil <- raster::crop(rast.soil,extent(list.rast.GCMs[[1]]))
rast.soil <- raster::resample(rast.soil,list.rast.GCMs[[1]])
rast.soil <- mask(rast.soil,PinpinDistri,updatevalue=NA)
  
# TRI raster
rast.tri <- raster("data/Topography/TRI_WGS84.tif")
rast.tri <- crop(rast.tri,extent(list.rast.GCMs[[1]]))
rast.tri <- raster::resample(rast.tri,list.rast.GCMs[[1]])
names(rast.tri) <- c("TRI")
rast.tri <- mask(rast.tri,PinpinDistri,updatevalue=NA)
  
# Burned Area
rast.BA <- raster("data/FireData/GFED4/PreparedRasters/GFED4_199507_201412_AllCountries.grd")
rast.BA <- crop(rast.BA,extent(list.rast.GCMs[[1]]))
rast.BA <- raster::resample(rast.BA,list.rast.GCMs[[1]])
names(rast.BA) <- "BurnedArea"
rast.BA <- mask(rast.BA,PinpinDistri,updatevalue=NA)
  

for(i in GCMs){
  
  stackall <- stack(list.rast.GCMs[[i]],rast.soil,rast.tri,rast.BA)
  stackall <- remove.NAs.stack(stackall) # same cells with NAs in all layers
  
  writeRaster(stackall, 
              filename=paste0("data/StacksEnvVars/FutureClimate/2041to2060/SSP585/Stack_",str_remove_all(i,"-"),".grd"), format="raster",overwrite=TRUE)
  
}
```

 
## Averaging GCMs predictions

We average the predictions of the nine (SSP3-7.0) or eight (SSP5-8.5) GCMs.


```{r FunctionAveragingGCMs}
# Function to average the predictions of the nine or eight GCMs

AverageGCMs <- function(SSP){

  path=paste0("data/StacksEnvVars/FutureClimate/2041to2060/",SSP,"/")
  file.names <- list.files(path=path,pattern="*grd$")

                    # Create a stack with nine layers (each corresponding to one GCM),
  # and in each layer, there is a stack with the 8 environmental variables
  stack.GCMs <- lapply(file.names, function(t) stack(paste0(path,t)))

  meanIgnoringZeroes <- function(x) {mean(x[x!=0],na.rm=T)}

  if(SSP=="SSP370"){
  AveragingClimGCMs <- function(clim.var,clim.stack){
    mean.clim <- overlay(
          clim.stack[[1]][[clim.var]],
          clim.stack[[2]][[clim.var]],
          clim.stack[[3]][[clim.var]],
          clim.stack[[4]][[clim.var]],
          clim.stack[[5]][[clim.var]],
          clim.stack[[6]][[clim.var]],
          clim.stack[[7]][[clim.var]],
          clim.stack[[8]][[clim.var]],
          clim.stack[[9]][[clim.var]],
          fun=meanIgnoringZeroes)
    return(mean.clim)
    
  }} else if(SSP=="SSP585"){
    
   AveragingClimGCMs <- function(clim.var,clim.stack){
    mean.clim <- overlay(
          clim.stack[[1]][[clim.var]],
          clim.stack[[2]][[clim.var]],
          clim.stack[[3]][[clim.var]],
          clim.stack[[4]][[clim.var]],
          clim.stack[[5]][[clim.var]],
          clim.stack[[6]][[clim.var]],
          clim.stack[[7]][[clim.var]],
          clim.stack[[8]][[clim.var]],
          fun=meanIgnoringZeroes)
    return(mean.clim)
  }   
    }

  avgbio5 <- AveragingClimGCMs(clim.var="bio5",clim.stack = stack.GCMs)
  avgbio6 <- AveragingClimGCMs(clim.var="bio6",clim.stack = stack.GCMs)
  avgbio12 <- AveragingClimGCMs(clim.var="bio12",clim.stack = stack.GCMs)
  avgbio15 <- AveragingClimGCMs(clim.var="bio15",clim.stack = stack.GCMs)


  stack.GCMs.avg <- stack(avgbio5,
                          avgbio6,
                          avgbio12,
                          avgbio15,
                          stack.GCMs[[1]][["water_top"]],
                          stack.GCMs[[1]][["depth_roots"]],
                          stack.GCMs[[1]][["TRI"]],
                          stack.GCMs[[1]][["BurnedArea"]])

  names(stack.GCMs.avg) <- c(clim.var,"water_top","depth_roots","TRI","BurnedArea")
  
  writeRaster(stack.GCMs.avg, 
              filename=paste0("data/StacksEnvVars/FutureClimate/2041to2060/",SSP,"/StackGCMsAvg.grd"),
              format="raster",
              overwrite=TRUE)
}
```


```{r ApplyAveragingGCMs}
SSPs <- c("SSP370","SSP585")

lapply(SSPs, AverageGCMs)
```




## Current climate (res 2.5 min)


```{r CurrentClimateRes25min}
rast.to.crop <- stack("data/StacksEnvVars/FutureClimate/2041to2060/SSP370/Stack_BCCCSM2MR.grd")

# Climatic rasters
# ----------------
grids <- list.files("data/climate/CurrentClimate/WorldClim30sec_1970_2000/")
grids <- grids[grepl("bio_5|bio_6|bio_12|bio_15",grids)==T]
grids <- c(grids[[3]],grids[[4]],grids[[1]],grids[[2]]) # reorder
rast.clim <- raster::stack(paste0("data/climate/CurrentClimate/WorldClim30sec_1970_2000/", grids))
names(rast.clim) <- clim.var
rast.clim <- crop(rast.clim,extent(rast.to.crop)) 
rast.clim <- raster::resample(rast.clim,rast.to.crop,method="ngb")
rast.clim <- mask(rast.clim,PinpinDistri, updatevalue=NA)


# Soil rasters
# ------------
rast.soil <- stack("data/soil/STU_EU_T_TAWC_WGS84.tif","data/soil/STU_EU_DEPTH_ROOTS_WGS84.tif")
names(rast.soil) <- c("water_top","depth_roots")
rast.soil <- raster::crop(rast.soil,extent(rast.clim))   # attribute the extent of climatic raster to soil raster
rast.soil <- raster::resample(rast.soil,rast.clim)       # same number of cells between the soil and climatic rasters
rast.soil <- mask(rast.soil,PinpinDistri,updatevalue=NA) # keep only cells within maritime pine distribution 


# TRI raster
# ----------
# > The models are fitted with TRI values from raster at 90m resolution, and the 
# projections will be on raster at 1km resolution (to have the same resolution as WorldClim variables).
rast.tri <- raster("data/Topography/TRI_WGS84.tif")
rast.tri <- crop(rast.tri,extent(rast.clim))           # attribute the extent of climatic raster to TRI rasters
rast.tri <- raster::resample(rast.tri,rast.clim)       # same number of cells between the TRI and climatic rasters
names(rast.tri) <- c("TRI")
rast.tri <- mask(rast.tri,PinpinDistri,updatevalue=NA) # keep only cells within maritime pine distribution 


# Burned Area raster
# ------------------
rast.BA <- raster("data/FireData/GFED4/PreparedRasters/GFED4_199507_201412_AllCountries.grd")
rast.BA <- crop(rast.BA,extent(rast.clim))
rast.BA <- raster::resample(rast.BA,rast.clim)
names(rast.BA) <- "BurnedArea"
rast.BA <- mask(rast.BA,PinpinDistri,updatevalue=NA)


# Merge and save the rasters in a stack
# ------------------------------------
stackall <- stack(rast.clim,rast.soil,rast.tri,rast.BA)
stackall <- remove.NAs.stack(stackall) # same cells with NAs in all layers
stackall
writeRaster(stackall, filename="data/StacksEnvVars/Stack_CurrentClimate_Res25minutes.grd", format="raster",overwrite=TRUE)
```

## Adjusting missing values

```{r AdjustingNAs}
stackCurrClimRes25min <- stack("data/StacksEnvVars/Stack_CurrentClimate_Res25minutes.grd")
stackFutClimSSP370 <- stack("data/StacksEnvVars/FutureClimate/2041to2060/SSP370/StackGCMsAvg.grd")
stackFutClimSSP585 <- stack("data/StacksEnvVars/FutureClimate/2041to2060/SSP585/StackGCMsAvg.grd")

stacks <- lapply(c(stackFutClimSSP370,stackFutClimSSP585), function(x){

cellval.fut <- na.omit(as.data.frame(raster::extract(x[[1]], PinpinDistri,cellnumbers=T)))
cellval.curr <- na.omit(as.data.frame(raster::extract(stackCurrClimRes25min[[1]], PinpinDistri,cellnumbers=T)))

# 15 cells to replace by NAs in the stack of future climate
cellval.fut[cellval.fut$cell %in% setdiff(cellval.fut$cell,cellval.curr$cell),]
cell1  <- cellval.fut$cell[cellval.fut$cell %in% setdiff(cellval.fut$cell,cellval.curr$cell)]

# 4 cells to replace by NAs in stackCurrClimRes25min
cellval.curr[cellval.curr$cell %in% setdiff(cellval.curr$cell,cellval.fut$cell),]
cell2 <- cellval.curr$cell[cellval.curr$cell %in% setdiff(cellval.curr$cell,cellval.fut$cell)]


for(i in 1:8)  {x[[i]][c(cell1)] <- NA}
for(i in 1:8)  {stackCurrClimRes25min[[i]][c(cell2)] <- NA}

list(x,stackCurrClimRes25min)
})


writeRaster(stacks[[1]][[2]], filename="data/StacksEnvVars/Stack_CurrentClimate_Res25minutes.grd", 
            format="raster",overwrite=TRUE)
writeRaster(stacks[[1]][[1]], 
              filename=paste0("data/StacksEnvVars/FutureClimate/2041to2060/SSP370/StackGCMsAvg.grd"),
              format="raster",
              overwrite=TRUE)
writeRaster(stacks[[2]][[1]], 
              filename=paste0("data/StacksEnvVars/FutureClimate/2041to2060/SSP585/StackGCMsAvg.grd"),
              format="raster",
              overwrite=TRUE)
```



`r knitr::opts_chunk$set(eval = F)`
