---
title: Projecting  under current and future climates the genomic variation and calculating
  the genomic offset
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
editor_options: 
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```


```{r setup, include=FALSE}
# knitr::opts_chunk$set(fig.width = 5,fig.height = 4,cache=TRUE,cache.lazy = FALSE)
knitr::opts_chunk$set(fig.width = 5,fig.height = 4,cache=FALSE)
options(width = 300)
library(knitr)
library(hierfstat)
library(reshape2)
library(ggbiplot)
library(gdm)
#library(cowplot)
#library(ggplot2)
#library(ggpubr)
#library(stringi)
library(beepr)
library(dplyr)
#library(devtools)
#library(xtable)
library(tidyverse)
#library(tibble)
library(gradientForest)
library(raster)
library(sgdm)
library(rgeos)
library(kableExtra)
```

```{r FuntionsOptionsUsed,echo=F}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

# Gray scale
grayscale_colors <- gray.colors(100,            # number of different color levels
                                start = 0.0,    # how black (0) to go
                                end = 1.0,      # how white (1) to go
                                gamma = 2.2,    # correction between how a digital
                                # camera sees the world and how human eyes see it
                                alpha = NULL)
```

# DATA

## Mapping

### Maritime pine distribution (mask)

We load the shapefile of the maritime pin distribution that comes from EUFORGEN and that we will use as a mask in the maps.

```{r LoadMaritimePineDistriMask}
PinpinDistri  <- shapefile('data/maps/MaskPinpinDistri/PinpinDistriEUforgen_NFIplotsBuffer10km.shp')
```

### Hillshade (background)

We load the hillshade background.

```{r LoadBackground}
background <- raster("data/maps/Hillshade/background.grd")
```

## Env data: AvgWater



```{r}
select.var.AvgWater <- c("bio1","bio2","bio12","bio14","water_top","depth_roots","TRI")
```


### Rasters of current climatic, soil and topographic variables

> Resolution ~ 1 km2 (30arc sec)

(0.93 x 0.93 = 0.86 km2 at the equator)

```{r LoadStackAvgWater}
stackCurrClimRes30sec <- stack("data/StacksEnvVars/CurrentClimate/Res30seconds/StackAvgWater.grd")
```

> Resolution ~ 30.869 km2 (2.5 minutes)

(5.556 x 5.556 at the equator)

```{r LoadStackAvgWaterLowRes}
stackCurrClimRes25min <- stack("data/StacksEnvVars/CurrentClimate/Res25min/StackAvgWater.grd")
```

### Rasters of future climatic, soil and topographic variables

```{r LoadFutureStackAvgWater}
# scenarion SSP370
path="data/StacksEnvVars/FutureClimate/2041to2060/SSP370/"
ssp370stack.names <- list.files(path=path,pattern="^StackAvgWater.*grd$")
ssp370stack <- lapply(ssp370stack.names, function(t) stack(paste0(path,t)))

# scenarion SSP585
path="data/StacksEnvVars/FutureClimate/2041to2060/SSP585/"
ssp585stack.names <- list.files(path=path,pattern="^StackAvgWater.*grd$")
ssp585stack <- lapply(ssp585stack.names, function(t) stack(paste0(path,t)))
```

```{r AvgStackGCMsSSP370, cache=T}
meanIgnoringZeroes <- function(x) {
  mean(x[x!=0],na.rm=T)}

AveragingClimGCMs <- function(clim.var,clim.stack){
  mean.clim <- overlay(
          clim.stack[[1]][[clim.var]],
          clim.stack[[2]][[clim.var]],
          clim.stack[[3]][[clim.var]],
          clim.stack[[4]][[clim.var]],
          clim.stack[[5]][[clim.var]],
          clim.stack[[6]][[clim.var]],
          clim.stack[[7]][[clim.var]],
          clim.stack[[8]][[clim.var]],
          clim.stack[[9]][[clim.var]],
          fun=meanIgnoringZeroes)
  return(mean.clim)
  }

avgbio1 <- AveragingClimGCMs(clim.var="bio1",clim.stack = ssp370stack)
avgbio2 <- AveragingClimGCMs(clim.var="bio2",clim.stack = ssp370stack)
avgbio12 <- AveragingClimGCMs(clim.var="bio12",clim.stack = ssp370stack)
avgbio14 <- AveragingClimGCMs(clim.var="bio14",clim.stack = ssp370stack)


ssp370stackavg <- stack(avgbio1,
                     avgbio2,
                     avgbio12,
                     avgbio14,
                     ssp370stack[[1]][["water_top"]],
                     ssp370stack[[1]][["depth_roots"]],
                     ssp370stack[[1]][["TRI"]])

names(ssp370stackavg) <- select.var.AvgWater
```


# Reference SNPs

```{r}
set="refSNPs"
```

## Projecting under current climate

### GDM

#### Projecting geo & env variables

```{r ProjGDMEnvGeoRefSNPs,cache=TRUE, fig.height=8, fig.width=8}
# Load the model
gdm.AvgWater <- readRDS(file=paste0("outputs/GDM/models/GDMAvgWater_",set,".rds"))

# Create a raster of the genetic importance of each environmental and geographical variable
rastTrans <- gdm.transform(gdm.AvgWater, stackCurrClimRes30sec) 

# Extract the values from the raster
rastDat <- na.omit(getValues(rastTrans))

# Principal Components Analysis (PCA) to reduce the transformed environmental variables into three factors
pca <- prcomp(rastDat, center=TRUE, scale.=FALSE)

##assigns to colors, edit as needed to maximize color contrast, etc.
a1 <- pca$x[,1]; a2 <- pca$x[,2]; a3 <- pca$x[,3]
r <- a1+a2; g <- -a2; b <- a3+a2-a1 

##scales colors
scalR <- (r-min(r))/(max(r)-min(r))*255
scalG <- (g-min(g))/(max(g)-min(g))*255
scalB <- (b-min(b))/(max(b)-min(b))*255

##assigns color to raster
rast1 <- rast2 <- rast3 <- stackCurrClimRes30sec[[1]]

rasValue <- raster::extract(stackCurrClimRes30sec, PinpinDistri,cellnumbers=T)
rasValue <- as.data.frame(rasValue)
rasValue <- na.omit(rasValue)


mapCells <- rasValue$cell
rast1[mapCells] <- scalR
rast2[mapCells] <- scalG
rast3[mapCells] <- scalB
##stacks color rasters
outRast <- stack(rast1, rast2, rast3)
plot(background,col=grayscale_colors, axes=FALSE,legend=F)
plotRGB(outRast, r = 1, g = 2, b = 3,add=T,bgalpha=0)

nvs <- dim(pca$rotation)[1]
vec <-  row.names(pca$rotation)
lv <- length(vec)
vind <- rownames(pca$rotation) %in% vec
scal <- 15
xrng <- range(pca$x[, 1], pca$rotation[, 1]/scal) *1.1
yrng <- range(pca$x[, 2], pca$rotation[, 2]/scal) * 1.1
pca$rotation
plot((pca$x[, 1:2]), xlim = xrng, ylim = yrng, pch = ".", cex = 4, col = rgb(scalR , scalG, scalB, max = 255), asp = 1)
points(pca$rotation[!vind, 1:2]/scal, pch = "+")
arrows(rep(0, lv), rep(0, lv), pca$rotation[vec,1]/scal, pca$rotation[vec, 2]/scal, length = 0.0625)
jit <- 0.0015
text(pca$rotation[vec, 1]/scal + jit * sign(pca$rotation[vec, 1]), 
     pca$rotation[vec, 2]/scal + jit * sign(pca$rotation[vec, 2]), 
     labels = vec)
```

#### Projecting only env variables

```{r ProjGDMEnvRefSNPs,cache=TRUE, fig.height=8, fig.width=8}
# Load the model
gdm.AvgWater <- readRDS(file=paste0("outputs/GDM/models/GDMAvgWater_",set,".rds"))

# Create a raster of the genetic importance of each environmental and geographical variable
rastTrans <- gdm.transform(gdm.AvgWater, stackCurrClimRes30sec) 

# Extract the values from the raster
rastDat <- na.omit(getValues(rastTrans))

# We may want to keep only the environmental variables, in this case, run the following line:
rastDat <- rastDat[,3:ncol(rastDat)]

# Principal Components Analysis (PCA) to reduce the transformed environmental variables into three factors
pca <- prcomp(rastDat, center=TRUE, scale.=FALSE)

##assigns to colors, edit as needed to maximize color contrast, etc.
a1 <- pca$x[,1]; a2 <- pca$x[,2]; a3 <- pca$x[,3]
r <- a1+a2; g <- -a2; b <- a3+a2-a1 

##scales colors
scalR <- (r-min(r))/(max(r)-min(r))*255
scalG <- (g-min(g))/(max(g)-min(g))*255
scalB <- (b-min(b))/(max(b)-min(b))*255

##assigns color to raster
rast1 <- rast2 <- rast3 <- stackCurrClimRes30sec[[1]]

rasValue <- raster::extract(stackCurrClimRes30sec, PinpinDistri,cellnumbers=T)
rasValue <- as.data.frame(rasValue)
rasValue <- na.omit(rasValue)


mapCells <- rasValue$cell
rast1[mapCells] <- scalR
rast2[mapCells] <- scalG
rast3[mapCells] <- scalB
##stacks color rasters
outRast <- stack(rast1, rast2, rast3)
plot(background,col=grayscale_colors, axes=FALSE,legend=F)
plotRGB(outRast, r = 1, g = 2, b = 3,add=T,bgalpha=0)

nvs <- dim(pca$rotation)[1]
vec <-  row.names(pca$rotation)
lv <- length(vec)
vind <- rownames(pca$rotation) %in% vec
scal <- 5
xrng <- range(pca$x[, 1], pca$rotation[, 1]/scal) *1.1
yrng <- range(pca$x[, 2], pca$rotation[, 2]/scal) * 1.1
pca$rotation
plot((pca$x[, 1:2]), xlim = xrng, ylim = yrng, pch = ".", cex = 4, col = rgb(scalR , scalG, scalB, max = 255), asp = 1)
points(pca$rotation[!vind, 1:2]/scal, pch = "+")
arrows(rep(0, lv), rep(0, lv), pca$rotation[vec,1]/scal, pca$rotation[vec, 2]/scal, length = 0.0625)
jit <- 0.0015
text(pca$rotation[vec, 1]/scal + jit * sign(pca$rotation[vec, 1]), 
     pca$rotation[vec, 2]/scal + jit * sign(pca$rotation[vec, 2]), 
     labels = vec)
```


## Projecting under future climate 

> scenario SSP370

### GDM

#### Projecting geo & env variables

```{r ProjGDMEnvGeoGCMsSSP370,fig.height=6,fig.width=6}
rastTrans <- gdm.transform(gdm.AvgWater, ssp370stackavg)
rastDat <-  na.omit(getValues(rastTrans))
rastDat <- as.data.frame(rastDat)

pca <- prcomp(rastDat, center=TRUE, scale.=FALSE) # bio12 does not appear as this variable has no explanatory power

##assigns to colors, edit as needed to maximize color contrast, etc.
a1 <- pca$x[,1]; a2 <- pca$x[,2]; a3 <- pca$x[,3]
r <- a1+a2; g <- -a2; b <- a3+a2-a1

##scales colors
scalR <- (r-min(r))/(max(r)-min(r))*255
scalG <- (g-min(g))/(max(g)-min(g))*255
scalB <- (b-min(b))/(max(b)-min(b))*255

##assigns color to raster
rast1 <- rast2 <- rast3 <- ssp370stackavg[[1]]

rasValue <- raster::extract(ssp370stackavg, PinpinDistri,cellnumbers=T)
rasValue <- as.data.frame(rasValue)
rasValue <- na.omit(rasValue)


mapCells <- rasValue$cell
rast1[mapCells] <- scalR
rast2[mapCells] <- scalG
rast3[mapCells] <- scalB

##stacks color rasters
outRast <- stack(rast1, rast2, rast3)
plot(background,col=grayscale_colors, axes=FALSE,legend=F)
plotRGB(outRast, r = 1, g = 2, b = 3,add=T,bgalpha=0)


nvs <- dim(pca$rotation)[1]
vec <-  row.names(pca$rotation)
lv <- length(vec)
vind <- rownames(pca$rotation) %in% vec
scal <- 5
xrng <- range(pca$x[, 1], pca$rotation[, 1]/scal) *1.1
yrng <- range(pca$x[, 2], pca$rotation[, 2]/scal) * 1.1
pca$rotation
plot((pca$x[, 1:2]), xlim = xrng, ylim = yrng, pch = ".", cex = 4, col = rgb(scalR , scalG, scalB, max = 255), asp = 1)
points(pca$rotation[!vind, 1:2]/scal, pch = "+")
arrows(rep(0, lv), rep(0, lv), pca$rotation[vec,1]/scal, pca$rotation[vec, 2]/scal, length = 0.0625)
jit <- 0.0015
text(pca$rotation[vec, 1]/scal + jit * sign(pca$rotation[vec, 1]), 
     pca$rotation[vec, 2]/scal + jit * sign(pca$rotation[vec, 2]), 
     labels = vec)
```

#### Projecting only env variables

```{r ProjGDMEnvGCMsSSP370,fig.height=6,fig.width=6}
rastTrans <- gdm.transform(gdm.AvgWater, ssp370stackavg)
rastDat <-  na.omit(getValues(rastTrans))
rastDat <- as.data.frame(rastDat)

# We may want to keep only the environmental variables, in this case, run the following line:
rastDat <- rastDat[,3:ncol(rastDat)]

pca <- prcomp(rastDat, center=TRUE, scale.=FALSE) # bio12 does not appear as this variable has no explanatory power

##assigns to colors, edit as needed to maximize color contrast, etc.
a1 <- pca$x[,1]; a2 <- pca$x[,2]; a3 <- pca$x[,3]
r <- a1+a2; g <- -a2; b <- a3+a2-a1

##scales colors
scalR <- (r-min(r))/(max(r)-min(r))*255
scalG <- (g-min(g))/(max(g)-min(g))*255
scalB <- (b-min(b))/(max(b)-min(b))*255

##assigns color to raster
rast1 <- rast2 <- rast3 <- ssp370stackavg[[1]]

rasValue <- raster::extract(ssp370stackavg, PinpinDistri,cellnumbers=T)
rasValue <- as.data.frame(rasValue)
rasValue <- na.omit(rasValue)


mapCells <- rasValue$cell
rast1[mapCells] <- scalR
rast2[mapCells] <- scalG
rast3[mapCells] <- scalB

##stacks color rasters
outRast <- stack(rast1, rast2, rast3)
plot(background,col=grayscale_colors, axes=FALSE,legend=F)
plotRGB(outRast, r = 1, g = 2, b = 3,add=T,bgalpha=0)


nvs <- dim(pca$rotation)[1]
vec <-  row.names(pca$rotation)
lv <- length(vec)
vind <- rownames(pca$rotation) %in% vec
scal <- 5
xrng <- range(pca$x[, 1], pca$rotation[, 1]/scal) *1.1
yrng <- range(pca$x[, 2], pca$rotation[, 2]/scal) * 1.1
pca$rotation
plot((pca$x[, 1:2]), xlim = xrng, ylim = yrng, pch = ".", cex = 4, col = rgb(scalR , scalG, scalB, max = 255), asp = 1)
points(pca$rotation[!vind, 1:2]/scal, pch = "+")
arrows(rep(0, lv), rep(0, lv), pca$rotation[vec,1]/scal, pca$rotation[vec, 2]/scal, length = 0.0625)
jit <- 0.0015
text(pca$rotation[vec, 1]/scal + jit * sign(pca$rotation[vec, 1]), 
     pca$rotation[vec, 2]/scal + jit * sign(pca$rotation[vec, 2]), 
     labels = vec)
```

# Common candidates

```{r}
set="CommonCandidates"
```

## Projecting under current climate

### GDM

#### Projecting geo & env variables

```{r ProjGDMEnvGeoComCandSNPs,cache=TRUE, fig.height=8, fig.width=8}
# Load the model
gdm.AvgWater <- readRDS(file=paste0("outputs/GDM/models/GDMAvgWater_",set,".rds"))

# Create a raster of the genetic importance of each environmental and geographical variable
rastTrans <- gdm.transform(gdm.AvgWater, stackCurrClimRes30sec) 

# Extract the values from the raster
rastDat <- na.omit(getValues(rastTrans))

# Principal Components Analysis (PCA) to reduce the transformed environmental variables into three factors
pca <- prcomp(rastDat, center=TRUE, scale.=FALSE)

##assigns to colors, edit as needed to maximize color contrast, etc.
a1 <- pca$x[,1]; a2 <- pca$x[,2]; a3 <- pca$x[,3]
r <- a1+a2; g <- -a2; b <- a3+a2-a1 

##scales colors
scalR <- (r-min(r))/(max(r)-min(r))*255
scalG <- (g-min(g))/(max(g)-min(g))*255
scalB <- (b-min(b))/(max(b)-min(b))*255

##assigns color to raster
rast1 <- rast2 <- rast3 <- stackCurrClimRes30sec[[1]]

rasValue <- raster::extract(stackCurrClimRes30sec, PinpinDistri,cellnumbers=T)
rasValue <- as.data.frame(rasValue)
rasValue <- na.omit(rasValue)


mapCells <- rasValue$cell
rast1[mapCells] <- scalR
rast2[mapCells] <- scalG
rast3[mapCells] <- scalB
##stacks color rasters
outRast <- stack(rast1, rast2, rast3)
plot(background,col=grayscale_colors, axes=FALSE,legend=F)
plotRGB(outRast, r = 1, g = 2, b = 3,add=T,bgalpha=0)

nvs <- dim(pca$rotation)[1]
vec <-  row.names(pca$rotation)
lv <- length(vec)
vind <- rownames(pca$rotation) %in% vec
scal <- 15
xrng <- range(pca$x[, 1], pca$rotation[, 1]/scal) *1.1
yrng <- range(pca$x[, 2], pca$rotation[, 2]/scal) * 1.1
pca$rotation
plot((pca$x[, 1:2]), xlim = xrng, ylim = yrng, pch = ".", cex = 4, col = rgb(scalR , scalG, scalB, max = 255), asp = 1)
points(pca$rotation[!vind, 1:2]/scal, pch = "+")
arrows(rep(0, lv), rep(0, lv), pca$rotation[vec,1]/scal, pca$rotation[vec, 2]/scal, length = 0.0625)
jit <- 0.0015
text(pca$rotation[vec, 1]/scal + jit * sign(pca$rotation[vec, 1]), 
     pca$rotation[vec, 2]/scal + jit * sign(pca$rotation[vec, 2]), 
     labels = vec)
```

#### Projecting only env variables

```{r ProjGDMEnvComCandSNPs,cache=TRUE, fig.height=8, fig.width=8}
# Load the model
gdm.AvgWater <- readRDS(file=paste0("outputs/GDM/models/GDMAvgWater_",set,".rds"))

# Create a raster of the genetic importance of each environmental and geographical variable
rastTrans <- gdm.transform(gdm.AvgWater, stackCurrClimRes30sec) 

# Extract the values from the raster
rastDat <- na.omit(getValues(rastTrans))

# We may want to keep only the environmental variables, in this case, run the following line:
rastDat <- rastDat[,3:ncol(rastDat)]

# Principal Components Analysis (PCA) to reduce the transformed environmental variables into three factors
pca <- prcomp(rastDat, center=TRUE, scale.=FALSE)

##assigns to colors, edit as needed to maximize color contrast, etc.
a1 <- pca$x[,1]; a2 <- pca$x[,2]; a3 <- pca$x[,3]
r <- a1+a2; g <- -a2; b <- a3+a2-a1 

##scales colors
scalR <- (r-min(r))/(max(r)-min(r))*255
scalG <- (g-min(g))/(max(g)-min(g))*255
scalB <- (b-min(b))/(max(b)-min(b))*255

##assigns color to raster
rast1 <- rast2 <- rast3 <- stackCurrClimRes30sec[[1]]

rasValue <- raster::extract(stackCurrClimRes30sec, PinpinDistri,cellnumbers=T)
rasValue <- as.data.frame(rasValue)
rasValue <- na.omit(rasValue)


mapCells <- rasValue$cell
rast1[mapCells] <- scalR
rast2[mapCells] <- scalG
rast3[mapCells] <- scalB
##stacks color rasters
outRast <- stack(rast1, rast2, rast3)
plot(background,col=grayscale_colors, axes=FALSE,legend=F)
plotRGB(outRast, r = 1, g = 2, b = 3,add=T,bgalpha=0)

nvs <- dim(pca$rotation)[1]
vec <-  row.names(pca$rotation)
lv <- length(vec)
vind <- rownames(pca$rotation) %in% vec
scal <- 5
xrng <- range(pca$x[, 1], pca$rotation[, 1]/scal) *1.1
yrng <- range(pca$x[, 2], pca$rotation[, 2]/scal) * 1.1
pca$rotation
plot((pca$x[, 1:2]), xlim = xrng, ylim = yrng, pch = ".", cex = 4, col = rgb(scalR , scalG, scalB, max = 255), asp = 1)
points(pca$rotation[!vind, 1:2]/scal, pch = "+")
arrows(rep(0, lv), rep(0, lv), pca$rotation[vec,1]/scal, pca$rotation[vec, 2]/scal, length = 0.0625)
jit <- 0.0015
text(pca$rotation[vec, 1]/scal + jit * sign(pca$rotation[vec, 1]), 
     pca$rotation[vec, 2]/scal + jit * sign(pca$rotation[vec, 2]), 
     labels = vec)
```



## Projecting under future climate 

> scenario SSP370

### GDM

#### Projecting geo & env variables

```{r ProjGDMEnvGeoGCMsSSP370ComCand,fig.height=6,fig.width=6}
rastTrans <- gdm.transform(gdm.AvgWater, ssp370stackavg)
rastDat <-  na.omit(getValues(rastTrans))
rastDat <- as.data.frame(rastDat)

pca <- prcomp(rastDat, center=TRUE, scale.=FALSE) # bio12 does not appear as this variable has no explanatory power

##assigns to colors, edit as needed to maximize color contrast, etc.
a1 <- pca$x[,1]; a2 <- pca$x[,2]; a3 <- pca$x[,3]
r <- a1+a2; g <- -a2; b <- a3+a2-a1

##scales colors
scalR <- (r-min(r))/(max(r)-min(r))*255
scalG <- (g-min(g))/(max(g)-min(g))*255
scalB <- (b-min(b))/(max(b)-min(b))*255

##assigns color to raster
rast1 <- rast2 <- rast3 <- ssp370stackavg[[1]]

rasValue <- raster::extract(ssp370stackavg, PinpinDistri,cellnumbers=T)
rasValue <- as.data.frame(rasValue)
rasValue <- na.omit(rasValue)


mapCells <- rasValue$cell
rast1[mapCells] <- scalR
rast2[mapCells] <- scalG
rast3[mapCells] <- scalB

##stacks color rasters
outRast <- stack(rast1, rast2, rast3)
plot(background,col=grayscale_colors, axes=FALSE,legend=F)
plotRGB(outRast, r = 1, g = 2, b = 3,add=T,bgalpha=0)


nvs <- dim(pca$rotation)[1]
vec <-  row.names(pca$rotation)
lv <- length(vec)
vind <- rownames(pca$rotation) %in% vec
scal <- 5
xrng <- range(pca$x[, 1], pca$rotation[, 1]/scal) *1.1
yrng <- range(pca$x[, 2], pca$rotation[, 2]/scal) * 1.1
pca$rotation
plot((pca$x[, 1:2]), xlim = xrng, ylim = yrng, pch = ".", cex = 4, col = rgb(scalR , scalG, scalB, max = 255), asp = 1)
points(pca$rotation[!vind, 1:2]/scal, pch = "+")
arrows(rep(0, lv), rep(0, lv), pca$rotation[vec,1]/scal, pca$rotation[vec, 2]/scal, length = 0.0625)
jit <- 0.0015
text(pca$rotation[vec, 1]/scal + jit * sign(pca$rotation[vec, 1]), 
     pca$rotation[vec, 2]/scal + jit * sign(pca$rotation[vec, 2]), 
     labels = vec)
```

#### Projecting only env variables

```{r ProjGDMEnvGCMsSSP370ComCand,fig.height=6,fig.width=6}
rastTrans <- gdm.transform(gdm.AvgWater, ssp370stackavg)
rastDat <-  na.omit(getValues(rastTrans))
rastDat <- as.data.frame(rastDat)

# We may want to keep only the environmental variables, in this case, run the following line:
rastDat <- rastDat[,3:ncol(rastDat)]

pca <- prcomp(rastDat, center=TRUE, scale.=FALSE) # bio12 does not appear as this variable has no explanatory power

##assigns to colors, edit as needed to maximize color contrast, etc.
a1 <- pca$x[,1]; a2 <- pca$x[,2]; a3 <- pca$x[,3]
r <- a1+a2; g <- -a2; b <- a3+a2-a1

##scales colors
scalR <- (r-min(r))/(max(r)-min(r))*255
scalG <- (g-min(g))/(max(g)-min(g))*255
scalB <- (b-min(b))/(max(b)-min(b))*255

##assigns color to raster
rast1 <- rast2 <- rast3 <- ssp370stackavg[[1]]

rasValue <- raster::extract(ssp370stackavg, PinpinDistri,cellnumbers=T)
rasValue <- as.data.frame(rasValue)
rasValue <- na.omit(rasValue)


mapCells <- rasValue$cell
rast1[mapCells] <- scalR
rast2[mapCells] <- scalG
rast3[mapCells] <- scalB

##stacks color rasters
outRast <- stack(rast1, rast2, rast3)
plot(background,col=grayscale_colors, axes=FALSE,legend=F)
plotRGB(outRast, r = 1, g = 2, b = 3,add=T,bgalpha=0)


nvs <- dim(pca$rotation)[1]
vec <-  row.names(pca$rotation)
lv <- length(vec)
vind <- rownames(pca$rotation) %in% vec
scal <- 5
xrng <- range(pca$x[, 1], pca$rotation[, 1]/scal) *1.1
yrng <- range(pca$x[, 2], pca$rotation[, 2]/scal) * 1.1
pca$rotation
plot((pca$x[, 1:2]), xlim = xrng, ylim = yrng, pch = ".", cex = 4, col = rgb(scalR , scalG, scalB, max = 255), asp = 1)
points(pca$rotation[!vind, 1:2]/scal, pch = "+")
arrows(rep(0, lv), rep(0, lv), pca$rotation[vec,1]/scal, pca$rotation[vec, 2]/scal, length = 0.0625)
jit <- 0.0015
text(pca$rotation[vec, 1]/scal + jit * sign(pca$rotation[vec, 1]), 
     pca$rotation[vec, 2]/scal + jit * sign(pca$rotation[vec, 2]), 
     labels = vec)
```


## Genomic offset

```{r SetSameNumberOfRowsRasters}
cellval.ssp370stack <- na.omit(as.data.frame(raster::extract(ssp370stack[[1]][[1]], PinpinDistri,cellnumbers=T)))
cellval.stackCurrClimRes25min <- na.omit(as.data.frame(raster::extract(stackCurrClimRes25min[[1]], PinpinDistri,cellnumbers=T)))

# 2 cells to replace by NAs in ssp370stack
cellval.ssp370stack[cellval.ssp370stack$cell %in% setdiff(cellval.ssp370stack$cell,cellval.stackCurrClimRes25min$cell),]
cell1  <- cellval.ssp370stack$cell[cellval.ssp370stack$cell %in% setdiff(cellval.ssp370stack$cell,cellval.stackCurrClimRes25min$cell)]

# 6 cells to replace by NAs in stackCurrClimRes25min
cellval.stackCurrClimRes25min[cellval.stackCurrClimRes25min$cell %in% setdiff(cellval.stackCurrClimRes25min$cell,cellval.ssp370stack$cell),]
cell2 <- cellval.stackCurrClimRes25min$cell[cellval.stackCurrClimRes25min$cell %in% setdiff(cellval.stackCurrClimRes25min$cell,cellval.ssp370stack$cell)]

replaceNA <- function(x) {for(i in 1:7)  {
  x[[i]][c(cell1)] <- NA
  return(x)}}

ssp370stack <- lapply(ssp370stack,replaceNA)
for(i in 1:7)  {stackCurrClimRes25min[[i]][c(cell2)] <- NA}
```


### GDM


```{r GOSSP370GDM, cache=TRUE, fig.height=3,fig.width=4.5}
GOAcrossRangeWithGDM <- function(x){
  
  rast.pred <- predict.gdm(gdm.AvgWater, stackCurrClimRes25min,time=TRUE, predRasts=x) 
  rasValue <- raster::extract(rast.pred, PinpinDistri,cellnumbers=T)
  rasValue <- as.data.frame(rasValue)
  rasValue <- na.omit(rasValue)
  return(rasValue$value)

} 

genOffallGCMs <- lapply(ssp370stack, GOAcrossRangeWithGDM)
names(genOffallGCMs) <- str_sub(ssp370stack.names,15,-5)

# convert to DF 
dfGOallGCMs <- as.data.frame(genOffallGCMs)

# Distribution of the genomic offset for each GCM
dfGOallGCMs %>%  
  pivot_longer(cols=everything(),names_to="model",values_to="go") %>% 
  ggplot(aes(x=go,color=model)) + 
    geom_density(size=1.5) +
    theme_bw() + 
    xlab("Genomic offset")


dfGOallGCMs$mean <- apply(dfGOallGCMs,1,mean)
dfGOallGCMs$sd <- apply(dfGOallGCMs,1,sd)
dfGOallGCMs$sdSup <- dfGOallGCMs$mean + dfGOallGCMs$sd
dfGOallGCMs$sdInf <- dfGOallGCMs$mean - dfGOallGCMs$sd
dfGOallGCMs$seSup <- dfGOallGCMs$mean + dfGOallGCMs$sd/sqrt(length(dfGOallGCMs$mean))
dfGOallGCMs$seInf <- dfGOallGCMs$mean - dfGOallGCMs$sd/sqrt(length(dfGOallGCMs$mean))


# Distribution of mean GO and its SE
dfGOallGCMs %>%  
  dplyr::select(mean,seSup,seInf) %>% 
  pivot_longer(cols=everything(),names_to="model",values_to="go") %>% 
  ggplot(aes(x=go,color=model)) + 
    geom_density(size=1.5) +
    theme_bw() + 
    xlab("Genomic offset")

# Distribution of mean GO and SD Supp and Inf
dfGOallGCMs %>%  
  dplyr::select(mean,sdSup,sdInf) %>% 
  pivot_longer(cols=everything(),names_to="model",values_to="go") %>% 
  ggplot(aes(x=go,color=model)) + 
    geom_density(size=1.5) +
    theme_bw() + 
    xlab("Genomic offset")

# Keeping mean and SD for futher analyses
#DFtotssp585 <- dfGOallGCMs %>% dplyr::select(mean,sd) %>% dplyr::rename(MeanBS=mean,SdBS=sd)
```

```{r PlotGOSSP370GDM,fig.height=8,fig.width=8}
rast.go <- ssp370stack[[1]][[1]] 
rasValue <- raster::extract(rast.go, PinpinDistri,cellnumbers=T)
rasValue <- as.data.frame(rasValue)
rasValue <- na.omit(rasValue)
rast.go[rasValue$cell] <- dfGOallGCMs$mean
# plot(rast.go,box = FALSE,
#      axes = FALSE,
#      alpha=0.8,
#       col=colorRampPalette(c("blue","yellow","red"))(255),
#      zlim=c(0,0.1))

plot(background,col=grayscale_colors, axes=FALSE,legend=F)
plot(rast.go,box = FALSE,
     axes = FALSE,
     #alpha=0.8,
      col=colorRampPalette(c("blue","yellow","red"))(255),
     #zlim=c(0,0.1),
     add=T,
     bgalpha=0)
```

Extract the Values of genomic offset for the NFI plots. 

```{r ExtractGOforNFIplotsGDM}
NFIdata <- readRDS(file="data/IFN/alexdata/dfplotPINPINA0.8R.M.rds") %>% 
  dplyr::select(plotcode,longitude,latitude,treeNbrJ.IMall) %>% 
  dplyr::rename(NbTot=treeNbrJ.IMall)
NFIdata <- NFIdata[!(NFIdata$NbTot==0),]
NFIdata$GO <- raster::extract(rast.go,NFIdata[,c("longitude","latitude")])

head(NFIdata) %>% 
  kable() %>%  
  kable_styling(bootstrap_options = c("striped","hover", "condensed"), full_width = F)

saveRDS(NFIdata,file="outputs/GDM/GO/nfiGO.rds")
```


`r knitr::opts_chunk$set(eval = F)`