---
title: "Validation in the common gardens"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5,fig.height = 4,cache=FALSE)
options(width = 300)
library(knitr)
library(tidybayes)
library(rstan)
library(brms)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
library(bayesplot)
color_scheme_set("green")
library(reshape2)
library(dplyr)
library(tidyverse)
library(xtable)
library(latex2exp)
library(kableExtra)
```

```{r CredibleIntervalsUsed, echo=F}
# The credible intervals used are 95% CI.
prob=0.95
probs <- c((1 - prob) / 2, 1 - (1 - prob) / 2)
prob_outer = 1
```


# Load CTD and genomic offset

```{r ExtractGOandCTD}
df <- lapply(c("GDM","GF"), function(x){
  
  df <- list.files(paste0("outputs/",x,"/ExtractedGenomicOffsetCG/"), full.names=TRUE) %>% 
    sapply(readRDS, USE.NAMES = TRUE,simplify=FALSE)
  names(df) <- str_sub(names(df),-7,-5)
  df <- df %>% 
    bind_rows(.id="SNPSets") %>% 
    mutate(Model=x)
  
}) %>% 
  bind_rows() %>% 
  mutate(CTD.bio1=s1.bio1-s2.bio1,
         CTD.bio5=s1.bio5-s2.bio5,
         CTD.bio6=s1.bio6-s2.bio6,
         CTD.bio12=s1.bio12-s2.bio12,
         CTD.bio15=s1.bio15-s2.bio15) %>% 
  pivot_wider(names_from=SNPSets,values_from = GO,names_prefix="GO.") %>% 
  pivot_wider(names_from=Model,values_from = c(GO.Ref,GO.Mer,GO.Mid,GO.Com),names_sep = ".") %>% 
  dplyr::select(-contains(c("s1.bio","s2.bio","depth","water","TRI","Burned","s1.x","s1.y"))) %>% 
  dplyr::rename(site=s2.site,
                prov=s1.prov)

df %>% 
  knitr::kable(digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"), full_width = F) %>%
  column_spec(1:2, bold = T)
```



# Mortality models

In this section, we want to determine whether **genomic offset (GO)** or **climate transfer distances (CTD)** are associated with the **proportion of dead trees in the populations**, independantly in two common gardens located in Spain (next to Madrid and Cáceres, respectively) and in which mortality rates were high. For that, we build a model that assumes that **tree height acts as a confounder**. Indeed, trees that were higher at the time of planting have a higher probability of survival. This is particularly true in Madrid and Cáceres where there were an extreme drought event the same year the trees were planted. Here is the model:
  
\begin{align*} 
a_{p} &\sim \text{Binomial} (N_{p},p_{p}) \\
\text{logit}(p_{p}) &= \beta_{0} +  \beta_{H}H_{p} + \beta_{X1}X_{p}\\
\end{align*}


with $a_{p}$ the count of individual that died in the population $p$, $N_{p}$ the total number of individuals in the population $p$ (=number of individuals that were initially planted in the common garden), $p_p$ is the estimated probability of mortality in the population $p$, $X_{p}$ is the genomic offset or climatic transfer distance for the population $p$ and $H_{p}$ is the BLUPs for height of the population $p$ (population varying intercepts calculated across all common gardens in the model 1 of \cite{archambeauCombiningClimaticGenomic2021}). We included $H_{p}$ as a covariate in the model to account for height differences before planting, as smaller trees had a higher mortality probability than taller trees. We used the following weakly informative priors:

  
\begin{align*} 
\begin{bmatrix}  \beta_{0,c} \\ \beta_{H} \\ \beta_{X_1} \end{bmatrix} &\sim \mathcal{N}(0,5)
\end{align*}



## Load the data

### Height data (confounder)

We use here the provenance intercepts obtained in the model 1 of the first paper. 

```{r ExtractHeightInterceptsPaper1Model1,message=F,warning=F}
HierMod <- readRDS(file="../../Pinpin_Clonapin/HeightPinpinClonapin/outputs/models/P1/MOD1.rds")
heights <- HierMod %>% broom::tidyMCMC(estimate.method = "mean",conf.int = T) %>% # we take the mean of the prov random intercepts
  filter(str_detect(term, "^(r_prov\\[)")) %>% 
  dplyr::rename(height=estimate,prov=term) %>% 
  mutate(prov=str_sub(prov,8,-12))

heights %>% 
  knitr::kable(digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"), full_width = F) %>%
  column_spec(1, bold = T)
```

### Mortality data (response variable)

```{r LoadSurvData}
data <- readRDS(file="data/ClonapinData/PhenoDataNovember2019_UpdatedSept2021_AllSites.rds")
```

```{r CheckIdentityDatasets, echoF, eval=F}
data.annual <- readRDS(file="data/ClonapinData/PhenoDataNovember2019_AnnualTraits_UpdatedSept2021_AllSites.rds")

TestIdentityDF <- function(sel.site,colum.site){
vec.annual <- data.annual[data.annual$site==sel.site,] %>%
    dplyr::select(prov,tree,all_of(colum.site)) %>%
    drop_na() %>% 
    arrange(tree)

vec.long <- data[data$site==sel.site,] %>%
    dplyr::select(prov,tree,survival) %>%
    drop_na() %>% 
    arrange(tree)

return(identical(pull(vec.annual[,3]),
                    vec.long$survival))
}


TestIdentityDF(sel.site= "caceres", colum.site ="CAC_survdec11")
TestIdentityDF(sel.site= "madrid", colum.site ="MAD_survdec11")
```




## Run the models

```{r CompileMortalityModel, cache=T,message=F,warning=F}
stancode = stan_model("scripts/StanModels/BinomialMortalityModelCG.stan")
print(stancode)
```


```{r RunMortalityModels, eval=F}
DF <- sapply(c("caceres","madrid"),function(sel.site){
  
vars <- colnames(df)[str_detect(colnames(df),"CTD|GO")]

counter <- 0

subdata <- data %>% 
  filter(site==sel.site) %>% 
  dplyr::select(prov,survival) %>% 
  drop_na() %>% 
  group_by(prov) %>% 
  dplyr::summarise(dead.count=n()-sum(survival),tot.count=n())

saveRDS(subdata, file=paste0("outputs/ValidationCG/ExperimentalDesignTables/Survival_",sel.site,".rds"))

subdf <- df %>% 
  filter(site==sel.site) %>% 
  inner_join(heights[,c("prov","height")], by="prov") %>% 
  inner_join(subdata, by="prov")

DFint <- sapply(vars,function(x){

  counter <<- counter + 1 
  
  X <- subdf[,x] %>% pull()

  stanlist <- list(N=length(subdf$prov),
                   NbDead=subdf$dead.count,
                   NbTot=subdf$tot.count,
                   H=(subdf$height-mean(subdf$height)/sd(subdf$height)),
                   X=(X -mean(X))/sd(X))

  # Running the model
  mstan <- sampling(stancode, data = stanlist, iter = 2000, chains = 4, cores = 4) 

  loo.mstan <- loo(mstan)
  saveRDS(loo.mstan,file=paste0("outputs/ValidationCG/Figs/MortalityModels/LOOs/loo_",vars[[counter]],"_",sel.site,".rds"))
  
  p <- mstan %>%  
    mcmc_areas(regex_pars = "beta", # "beta|alpha"
                        prob=prob,
                        prob_outer=prob_outer,
                        point_est = "median") +  
    theme_bw() + 
    scale_y_discrete(labels=c("beta0"=parse(text=TeX("$\\beta_{0}$")),
                              "betaH"=parse(text=TeX("$\\beta_{H}$")),
                              'betaX1'=parse(text = TeX("$\\beta_{X1}$")),
                              'betaX2'=parse(text = TeX("$\\beta_{X2}$")))) +
    ggtitle(paste0(vars[[counter]]," - ", sel.site)) +
    theme(axis.text = element_text(size=16)) 
  
  ggsave(p,file=paste0("outputs/ValidationCG/Figs/MortalityModels/MCMCareas/MCMCareas_",vars[[counter]],"_",sel.site,".pdf"),device="pdf")
  
  
  # Save coefficients
  broom::tidyMCMC(mstan,
                  droppars = NULL, 
                  estimate.method = "median", 
                  ess = F, 
                  rhat = F, 
                  conf.int = T,
                  conf.level = prob) %>% 
  filter(str_detect(term, c('beta')))

}, USE.NAMES = TRUE,simplify=FALSE) %>% 
  
  bind_rows(.id="variable")
  
}, USE.NAMES = TRUE,simplify=FALSE) %>% 
  
  bind_rows(.id="site")

saveRDS(DF,file="outputs/ValidationCG/TablesPosteriors/OutputsMortalityModels.rds")
```

```{r LookTablePosteriorsMortalityModels}
DF <- readRDS(file="outputs/ValidationCG/TablesPosteriors/OutputsMortalityModels.rds")
DF %>% 
  print(n=Inf) %>%
  knitr::kable(digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"), full_width = F)
```

## Visualization

```{r SummaryFigsMortalityModels}
index.match <- list(betaH=c("betaH","Height effect ($\\beta_{H}$ estimate)"),
                    betaX1=c("betaX1","Regression coefficients"))

lapply(c("betaX1","betaH"),function(index){
  
p <- readRDS(file=paste0("outputs/ValidationCG/TablesPosteriors/OutputsMortalityModels.rds")) %>% 
  filter(term==index) %>% 
  mutate(Method=case_when(str_detect(variable,"CTD")==TRUE~"Climatic transfer distance",
                          str_detect(variable,"GDM")==TRUE~"Genomic offset",
                          str_detect(variable,"GF")==TRUE~"Genomic offset"),
         Method.GO=case_when(str_detect(variable,"CTD")==TRUE~"CTD",
                             str_detect(variable,"GDM")==TRUE~"GDM",
                          str_detect(variable,"GF")==TRUE~"GF"),
         Variable=case_when(str_detect(variable,"CTD")==TRUE~str_sub(variable,5,-1),
                            str_detect(variable,"GO")==TRUE~str_sub(variable,4,6)),
         Variable=factor(Variable,levels=c(paste0("bio",c(1,5,6,12,15)),"Com","Mid","Mer","Ref")),
         site=case_when(site=="caceres"~"Cáceres",
                        site=="madrid"~"Madrid")) %>% 
  
  ggplot(aes(x = Variable, y = estimate,ymin = conf.low, ymax = conf.high,colour=Method.GO,shape=Method.GO)) +#
  geom_hline(yintercept = 0,color="gray") +
  geom_pointinterval(position = position_dodge(width = .4),point_size=3,size=2,show.legend = c(size = TRUE)) +
  facet_grid(site~Method,scales="free_x", space = "free") + 
  ylab(TeX(index.match[[index]][[2]])) + xlab("") +
  scale_colour_manual(values=c("mediumpurple1","turquoise2","forestgreen"),
                      breaks = c('GF', 'GDM')) +
  scale_shape_manual(values = c(15, 16,17),
                     breaks = c('GF', 'GDM')) +
  theme_bw() +
  labs(color="Genomic offset approach:",shape="Genomic offset approach:") +
  theme(axis.text.x = element_text(size=13),
        axis.text.y = element_text(size=13),
        axis.title = element_text(size=16),
      legend.title=element_text(size=13), 
      legend.text=element_text(size=10),
      legend.position = "bottom",
      legend.background = element_rect(colour = "grey"),
      strip.text.x = element_text(size = 16),
      strip.text.y = element_text(size=16,angle = 0),
      strip.background = element_blank(),
      panel.grid.minor.x=element_blank(),
      panel.grid.major.x=element_blank()) +
    guides(color=guide_legend(ncol=2))
  
# for latex
ggsave(p,file=paste0("outputs/ValidationCG/Figs/MortalityModels/SummaryFigures/",index.match[[index]][[1]],".pdf"),
       device="pdf",
       height=9,
       width=10)

# for google docs
ggsave(p,file=paste0("outputs/ValidationCG/Figs/MortalityModels/SummaryFigures/",index.match[[index]][[1]],".png"),
       height=9,
       width=10)
})
```


## Leave-one out cross validation

```{r LooMortalityModels}
lapply(c("caceres","madrid"),function(sel.site){
  
d <- list.files(paste0("outputs/ValidationCG/Figs/MortalityModels/LOOs"),full.names=TRUE) %>% 
  str_subset(sel.site) %>% 
  sapply(readRDS, USE.NAMES = TRUE,simplify=FALSE) %>% 
  loo_compare() %>% 
  as.data.frame() %>% 
  rownames_to_column(var ="variable") %>% 
  as_tibble() %>% 
  mutate(site=sel.site,
         variable=str_sub(variable,52,-12),
         ModellingApproach=case_when(str_detect(variable,"GF")~"GF-based genomic offset",
                                     str_detect(variable,"GDM")~"GDM-based genomic offset",
                                     str_detect(variable,"CTD")~"Climatic transfer distance"),
         Covariate=case_when(str_detect(variable,"Mid")~"Candidates under expected strong selection",
                             str_detect(variable,"Mer")~"Merged candidates",
                             str_detect(variable,"Com")~"Common candidates",
                             str_detect(variable,"Ref")~"Reference candidates",
                             str_detect(variable,"bio5")~"bio5 - Max T° of the warmest month (°C)",
                             str_detect(variable,"bio6")~"bio6 - Min T° of the coldest month (°C)",
                             str_detect(variable,"bio12")~"bio12 - Annual precipitation (mm)",
                             str_detect(variable,"bio15")~"bio15 - Precipitation seasonality (coeff of variation)",
                             str_detect(variable,"bio1")~"bio1 - Annual daily mean T° (°C)")) %>% 
    dplyr::select(ModellingApproach,Covariate,elpd_diff,se_diff)

print(xtable(d,type = "latex",digits=2), 
      file = paste0("outputs/ValidationCG/TablesLOOs/MortalityLoos_",sel.site,".tex"), 
      include.rownames=FALSE)

d
  
})
```

**How to interpret the outputs from `loo_compare` function?**

- [How to use cross-validation for model selection? in the Cross-validation FAQ of Aki Vehtari](https://avehtari.github.io/modelselection/CV-FAQ.html#how-to-use-cross-validation-for-model-selection)


"If there is no application specific utility or loss function, use log score, i.e. elpd. **If elpd difference** (`elpd_diff` in `loo` package) **is less than 4, the difference is small**. If elpd difference is larger than 4, then compare that difference to standard error of `elpd_diff` (provided e.g. by `loo` package)."

- [Interpreting output from compare() of loo](https://discourse.mc-stan.org/t/interpreting-output-from-compare-of-loo/3380/2)


"it’s better to consider something like 2SE or more cautious 4SE, where 4 comes from the fact that SE for LOO can be underestimated for small n or under bad model misspecification.""

## Experimental design

We generate a latex table for the Supplementary Information.

```{r ExpDesignMortalityModels}
ExpDesignTab <- sapply(c("caceres","madrid"),function(sel.site){
  
  readRDS(file=paste0("outputs/ValidationCG/ExperimentalDesignTables/Survival_",sel.site,".rds")) %>% 
    mutate(prop.dead=dead.count*100/tot.count)
  
}, USE.NAMES = TRUE,simplify=FALSE ) %>% 
  bind_rows(.id="site") %>% 
  pivot_wider(names_from=site,values_from = c(dead.count, tot.count,prop.dead),names_sep="_") %>% 
  dplyr::select(prov,contains("caceres"),contains("madrid"))

# Generate the latex table
print(xtable(ExpDesignTab, type = "latex",digits=2), 
      file = paste0("outputs/ValidationCG/ExperimentalDesignTables/MortalityData.tex"), 
      include.rownames=FALSE)

# Information used in the manuscript
ExpDesignTab[,-1] %>% 
  dplyr::summarise_all(mean) %>% 
  knitr::kable(digits = 2) %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"), full_width = F)
```



# Height models

## Mathematical model

In this section, we want to determine whether **genomic offset (GO)** or **climate transfer distances (CTD)** are associated with the **mean height of the populations**, independantly in five common gardens located in Spain (Asturias, Madrid, Cáceres), Portugal (near Fundao) and France (Pierroton). In each common garden $s$, we perfom the following model:
  
\begin{align*} 
H_{p}  &\sim \mathcal{N}(\mu_{p},\sigma^{2}_{r}) \\
\mu_{p} &= \beta_{0} + \beta_{X1}X_{p} + \beta_{X2}X^{2}_{p} \\
\end{align*}

with $H_{p}$ the BLUPs for height of the population $p$ in the site $s$ (calculated with site-specific models in which the effects of blocks and the genetic population structure were accounted for, see script `EstimatingBLUPsForHeight.R`) and $X_{p}$ is the value of the variable of interest (GO or CTD) in the population $p$. We include a quadratic term for $X_p$ to allow for potential nonlinearity in the response, following @fitzpatrick2021experimental.

<!-- First version: The BLUPs were calculated with the model 2 of [Archambeau et al. 2021](https://www.biorxiv.org/content/10.1101/2020.11.13.382515v1.abstract), which accounts for the population-by-site interaction -->


```{r CompileHeightModel, cache=T,message=F,warning=F}
stancode = stan_model("scripts/StanModels/GaussianModelHeightCG.stan")
print(stancode)
```

## Running the models

<!-- ## Using intercepts from paper 1 -->

<!-- We use here the provenance intercepts obtained in the model 1 of the first paper.  -->

```{r ExtractHeightInterceptsPaper1Model2orModel4,message=F,warning=F,eval=F,echo=F}
heights <- readRDS(file="../../Pinpin_Clonapin/HeightPinpinClonapin/outputs/models/P1/MOD2.rds") %>% 
  broom::tidyMCMC(estimate.method = "mean",conf.int = T) %>%   # we take the mean of the prov random intercepts
  filter(str_detect(term, "^(r_prov:site)|^(r_prov\\[)")) 

heights <- readRDS(file="../../Pinpin_Clonapin/HeightPinpinClonapin/outputs/models/P1/MOD4.rds") %>% 
  broom::tidyMCMC(estimate.method = "mean",conf.int = T) %>%   # we take the mean of the prov random intercepts
  filter(str_detect(term, "^(r_prov:site)|^(r_prov\\[)")) 
  # dplyr::rename(height=estimate,prov=term) %>% 
  # mutate(prov=str_sub(prov,8,-12))

heights %>% 
  knitr::kable(digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"), full_width = F) %>%
  column_spec(1, bold = T)

sites <- c("caceres","madrid","asturias","bordeaux","portugal")

DF <- sapply(sites,function(sel.site){

site.intercepts <- heights %>%
  filter(term %in% term[str_detect(term,"^(r_prov:site\\[)")]) %>%
  filter(term %in% term[str_detect(term,sel.site)]) %>%
  mutate(prov=str_sub(term,13,15)) %>%
  dplyr::rename(site.estimate=estimate) %>%
  dplyr::select(prov,site.estimate)

intercepts <- heights %>%
  filter(term %in% term[str_detect(term,"^(r_prov\\[)")]) %>%
  mutate(prov=str_sub(term,8,-12)) %>%
  dplyr::rename(mean.estimate=estimate) %>%
  dplyr::select(prov,mean.estimate) %>%
  inner_join(site.intercepts,by="prov") %>%
  mutate(estimate=mean.estimate+site.estimate)

  
vars <- colnames(df)[str_detect(colnames(df),"CTD|GO")]

counter <- 0

DFint <- sapply(vars,function(x){

  counter <<- counter + 1 
  
  subdf <- df %>% 
    filter(site==sel.site) %>% 
    inner_join(intercepts[,c("prov","estimate")], by="prov")

  
  X <- subdf[,x] %>% pull()

  stanlist <- list(N=length(subdf$prov),
                   H=subdf$estimate,
                   X=(X -mean(X))/sd(X))

  # Running the model
  mstan <- sampling(stancode, data = stanlist, iter = 2000, chains = 4, cores = 4) 
  
  loo.mstan <- loo(mstan)
  saveRDS(loo.mstan,file=paste0("outputs/ValidationCG/Figs/HeightModels/LOOs/loo_",vars[[counter]],"_",sel.site,".rds"))

  p <- mstan %>%  
    mcmc_areas(regex_pars = "beta", # "beta|alpha"
                        prob=prob,
                        prob_outer=prob_outer,
                        point_est = "median") +  
    theme_bw() + 
    scale_y_discrete(labels=c("beta0"=parse(text=TeX("$\\beta_{0}$")),
                              "betaH"=parse(text=TeX("$\\beta_{H}$")),
                              'betaX1'=parse(text = TeX("$\\beta_{X1}$")),
                              'betaX2'=parse(text = TeX("$\\beta_{X2}$")))) +
    ggtitle(paste0(vars[[counter]]," - ", sel.site)) +
    theme(axis.text = element_text(size=16)) 
  
  #ggsave(p,file=paste0("outputs/ValidationCG/Figs/HeightModels/MCMCareas_",vars[[counter]],"_",sel.site,".pdf"),device="pdf")
  
  
  # Save coefficients
  d <- broom::tidyMCMC(mstan,
                  droppars = NULL, 
                  estimate.method = "median", 
                  ess = F, 
                  rhat = F, 
                  conf.int = T,
                  conf.level = prob) %>% 
  filter(str_detect(term, c('R_squared|beta')))
  
  rm(mstan)
  
  d

}, USE.NAMES = TRUE,simplify=FALSE) %>% 
  
  bind_rows(.id="variable")
  
}, USE.NAMES = TRUE,simplify=FALSE) %>% 
  
  bind_rows(.id="site")


```

```{r AccountingOrNotForPopStructure}
# !!! Do we account for population structure or not in the BLUPs? !!!
PopStructure <- "NotAccountingPopStructure" # AccountingPopStructure or NotAccountingPopStructure
```


```{r RunHeightModels, eval=F}
# List of height measurements we look at:
sites <- c("MAD_htdec11","CAC_htdec11","BDX_htnov13","BDX_htnov18","POR_htjan12","POR_htmay13","AST_htdec11","AST_htmar14")

match.site <- list(MAD_htdec11="madrid",
              CAC_htdec11="caceres",
              BDX_htnov13="bordeaux",
              BDX_htnov18="bordeaux",
              POR_htjan12="portugal",
              POR_htmay13="portugal",
              AST_htdec11="asturias",
              AST_htmar14="asturias")

# List model files
mod.files <-  list.files(paste0("outputs/ValidationCG/ModelsHeightBLUPs/",PopStructure,"/Models"), full.names=TRUE)

# Launch the function to run one model per height measurement:
DF <- sapply(sites,function(sel.site){
 
# Extract the BLUPs (mean intercept of each population)
intercepts  <- readRDS(file=paste0("outputs/ValidationCG/ModelsHeightBLUPs/",PopStructure,"/Models/",sel.site,".rds")) %>% 
  broom::tidyMCMC(estimate.method = "mean",conf.int = T) %>%   # we take the mean of the prov random intercepts
  filter(str_detect(term, "^(r_prov\\[)")) %>% 
  mutate(prov=str_sub(term,8,-12)) %>% 
  dplyr::select(prov,estimate)
  
vars <- colnames(df)[str_detect(colnames(df),"CTD|GO")]

# Merge the mean population intercepts with the variables of interest (CTD and GO)
subdf <- df %>% 
    filter(site==match.site[[sel.site]]) %>% 
    inner_join(intercepts[,c("prov","estimate")], by="prov")
  
counter <- 0

DFint <- sapply(vars,function(x){

  counter <<- counter + 1 
  
  X <- subdf[,x] %>% pull()
  
  # List for Stan
  stanlist <- list(N=length(subdf$prov),
                   H=subdf$estimate,
                   X=(X -mean(X))/sd(X))

  # Running the model with rstan
  mstan <- sampling(stancode, data = stanlist, iter = 2000, chains = 4, cores = 4) 
  
  # Calculating  the LOO to compare the models
  loo.mstan <- loo(mstan)
  saveRDS(loo.mstan,file=paste0("outputs/ValidationCG/Figs/HeightModels/LOOs/loo_",vars[[counter]],"_",sel.site,".rds"))

  # Plot the posteriors of the regression coefficients (beta_0, beta_X1 and beta_X2) and R^2
  p <- mstan %>%  
    mcmc_areas(regex_pars = "beta", # "beta|alpha"
                        prob=prob,
                        prob_outer=prob_outer,
                        point_est = "median") +  
    theme_bw() + 
    scale_y_discrete(labels=c("beta0"=parse(text=TeX("$\\beta_{0}$")),
                              "betaH"=parse(text=TeX("$\\beta_{H}$")),
                              'betaX1'=parse(text = TeX("$\\beta_{X1}$")),
                              'betaX2'=parse(text = TeX("$\\beta_{X2}$")))) +
    ggtitle(paste0(vars[[counter]]," - ", sel.site)) +
    theme(axis.text = element_text(size=16)) 
  
  ggsave(p,file=paste0("outputs/ValidationCG/Figs/HeightModels/",PopStructure,"/MCMCareas_",vars[[counter]],"_",sel.site,".pdf"),device="pdf")
  
  
  # Save the coefficients in a table
  d <- broom::tidyMCMC(mstan,
                  droppars = NULL, 
                  estimate.method = "median", 
                  ess = F, 
                  rhat = F, 
                  conf.int = T,
                  conf.level = prob) %>% 
  filter(str_detect(term, c('R_squared|beta')))
  
  rm(mstan)
  
  d

}, USE.NAMES = TRUE,simplify=FALSE) %>% 
  
  bind_rows(.id="variable")
  
}, USE.NAMES = TRUE,simplify=FALSE) %>% 
  
  bind_rows(.id="site")

saveRDS(DF,file=paste0("outputs/ValidationCG/TablesPosteriors/OutputsHeightModels_",PopStructure,".rds"))
```

```{r LookTablePosteriorsHeightModels, eval=F}
DF <- readRDS(file=paste0("outputs/ValidationCG/TablesPosteriors/OutputsHeightModels_",PopStructure,".rds"))

DF %>% 
  filter(term=="betaX1"|term=="betaX2") %>% 
  print(n=Inf) %>%
  knitr::kable(digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"), full_width = F)

DF %>% 
  filter(term=="R_squared") %>% 
  print(n=Inf) 

DF %>% 
  filter(term=="betaX1") %>% 
  print(n=Inf)

```

## Visualization

```{r SummaryFigsHeightModels,fig.height=12,fig.width=12}
index.match <- list(R_squared=c("R2","Goodness of fit ($R^{2}$ estimate)"),
                    betaX1=c("betaX1","Regression coefficients"))

lapply(c("betaX1","R_squared"),function(index){


pdf <- readRDS(file=paste0("outputs/ValidationCG/TablesPosteriors/OutputsHeightModels_",PopStructure,".rds")) %>% 
  filter(term==index) %>% 
  mutate(Method=case_when(str_detect(variable,"CTD")==TRUE~"Climatic transfer distance",
                          str_detect(variable,"GDM")==TRUE~"Genomic offset",
                          str_detect(variable,"GF")==TRUE~"Genomic offset"),
         Method.GO=case_when(str_detect(variable,"CTD")==TRUE~"CTD",
                             str_detect(variable,"GDM")==TRUE~"GDM",
                          str_detect(variable,"GF")==TRUE~"GF"),
         Variable=case_when(str_detect(variable,"CTD")==TRUE~str_sub(variable,5,-1),
                            str_detect(variable,"GO")==TRUE~str_sub(variable,4,6)),
         Variable=factor(Variable,levels=c(paste0("bio",c(1,5,6,12,15)),"Com","Mid","Mer","Ref")),
         Site.labels=case_when(str_detect(site,"MAD")~"Madrid",
                               str_detect(site,"CAC")~"Cáceres",
                               str_detect(site,"BDX")~"Bordeaux",
                               str_detect(site,"POR")~"Portugal",
                               str_detect(site,"AST")~"Asturias"),
         Age=case_when(site=="MAD_htdec11"~ "(13 months)",
                       site=="CAC_htdec11"~ "(8 months)",
                       site=="BDX_htnov13"~ "(25 months)",
                       site=="BDX_htnov18"~ "(85 months)",
                       site=="POR_htjan12"~ "(11 months old)",
                       site=="POR_htmay13"~ "(27 months old)",
                       site=="AST_htdec11"~ "(10 months old)",
                       site=="AST_htmar14"~ "(37 months old)"),
         
         Site.Age.labels=case_when(site=="MAD_htdec11"~ "Madrid (13 months)",
                                   site=="CAC_htdec11"~ "Cáceres (8 months)",
                                   site=="BDX_htnov13"~ "Bordeaux (25 months)",
                                   site=="BDX_htnov18"~ "Bordeaux (85 months)",
                                   site=="POR_htjan12"~ "Portugal (11 months old)",
                                   site=="POR_htmay13"~ "Portugal (27 months old)",
                                   site=="AST_htdec11"~ "Asturias (10 months old)",
                                   site=="AST_htmar14"~ "Asturias (37 months old)")) 

# All combinations site/age: 
p <- pdf %>% 
  ggplot(aes(x = Variable, y = estimate,ymin = conf.low, ymax = conf.high,colour=Method.GO,shape=Method.GO)) +#
  geom_hline(yintercept = 0,color="gray") +
  geom_pointinterval(position = position_dodge(width = .4),point_size=3,size=2,show.legend = c(size = TRUE)) +
  facet_grid(Site.Age.labels~Method,scales="free_x", space = "free") + 
  ylab(TeX(index.match[[index]][[2]])) + xlab("") +
  scale_colour_manual(values=c("mediumpurple1","turquoise2","forestgreen"),
                      breaks = c('GF', 'GDM')) +
  scale_shape_manual(values = c(15, 16,17),
                     breaks = c('GF', 'GDM')) +
  theme_bw() +
  labs(color="Genomic offset approach:",shape="Genomic offset approach:") +
  theme(axis.text.x = element_text(size=13),
        axis.text.y = element_text(size=13),
        axis.title = element_text(size=16),
      legend.title=element_text(size=13), 
      legend.text=element_text(size=10),
      legend.position = "bottom",
      legend.background = element_rect(colour = "grey"),
      strip.text.x = element_text(size = 14),
      strip.text.y = element_text(size=12,angle = 0),
      panel.grid.minor.x=element_blank(),
      panel.grid.major.x=element_blank(),
      strip.background = element_blank()) +
    guides(color=guide_legend(ncol=2))
  
ggsave(p,
       file=paste0("outputs/ValidationCG/Figs/HeightModels/SummaryFigures/",index.match[[index]][[1]],"_",PopStructure,".pdf"),
       device="pdf",
       width=9,
       height=15)

# One age per site:
p <- pdf %>% 
  filter(!(site %in% c("BDX_htnov13","POR_htjan12","AST_htdec11"))) %>%  
  ggplot(aes(x = Variable, y = estimate,ymin = conf.low, ymax = conf.high,colour=Method.GO,shape=Method.GO)) +#
  geom_hline(yintercept = 0,color="gray") +
  geom_pointinterval(position = position_dodge(width = .4),point_size=3,size=2,show.legend = c(size = TRUE)) +
  facet_grid(Site.labels~Method,scales="free_x", space = "free") + 
  ylab(TeX(index.match[[index]][[2]])) + xlab("") +
  scale_colour_manual(values=c("mediumpurple1","turquoise2","forestgreen"),
                      breaks = c('GF', 'GDM')) +
  scale_shape_manual(values = c(15, 16,17),
                     breaks = c('GF', 'GDM')) +
  theme_bw() +
  labs(color="Genomic offset approach:",shape="Genomic offset approach:") +
  theme(axis.text.x = element_text(size=13),
        axis.text.y = element_text(size=13),
        axis.title = element_text(size=16),
      legend.title=element_text(size=13), 
      legend.text=element_text(size=10),
      legend.position = "bottom",
      legend.background = element_rect(colour = "grey"),
      strip.text.x = element_text(size = 14),
      strip.text.y = element_text(size=12,angle = 0),
      panel.grid.minor.x=element_blank(),
      panel.grid.major.x=element_blank(),
      strip.background = element_blank()) +
    guides(color=guide_legend(ncol=2))

# for latex
ggsave(p,
       file=paste0("outputs/ValidationCG/Figs/HeightModels/SummaryFigures/",index.match[[index]][[1]],"_",PopStructure,"_Reduced.pdf"),
       device="pdf",
       width=9,
       height=12)

# for google docs:
ggsave(p,
       file=paste0("outputs/ValidationCG/Figs/HeightModels/SummaryFigures/",index.match[[index]][[1]],"_",PopStructure,"_Reduced.png"),
       width=9,
       height=12)
  

})

```

## Leave-one-out cross validation


```{r LooHeightModels}
sites <- c("MAD_htdec11","CAC_htdec11","BDX_htnov13","BDX_htnov18","POR_htjan12","POR_htmay13","AST_htdec11","AST_htmar14")

sapply(sites,function(sel.site){
  
d <- list.files(paste0("outputs/ValidationCG/Figs/HeightModels/LOOs"),full.names=TRUE) %>% 
  str_subset(sel.site) %>% 
  sapply(readRDS, USE.NAMES = TRUE,simplify=FALSE) %>% 
  loo_compare() %>% 
  as.data.frame() %>% 
  rownames_to_column(var ="variable") %>% 
  as_tibble() %>% 
  mutate(site=sel.site,
         variable=str_sub(variable,49,-17),
         ModellingApproach=case_when(str_detect(variable,"GF")~"GF-based genomic offset",
                                     str_detect(variable,"GDM")~"GDM-based genomic offset",
                                     str_detect(variable,"CTD")~"Climatic transfer distance"),
         Covariate=case_when(str_detect(variable,"Mid")~"Candidates under expected strong selection",
                             str_detect(variable,"Mer")~"Merged candidates",
                             str_detect(variable,"Com")~"Common candidates",
                             str_detect(variable,"Ref")~"Reference candidates",
                             str_detect(variable,"bio5")~"bio5 - Max T° of the warmest month (°C)",
                             str_detect(variable,"bio6")~"bio6 - Min T° of the coldest month (°C)",
                             str_detect(variable,"bio12")~"bio12 - Annual precipitation (mm)",
                             str_detect(variable,"bio15")~"bio15 - Precipitation seasonality (coeff of variation)",
                             str_detect(variable,"bio1")~"bio1 - Annual daily mean T° (°C)")) %>% 
    dplyr::select(ModellingApproach,Covariate,elpd_diff,se_diff)

print(xtable(d,type = "latex",digits=2), 
      file = paste0("outputs/ValidationCG/TablesLOOs/HeightLoos_",sel.site,".tex"), 
      include.rownames=FALSE)

d
  
}, USE.NAMES = TRUE,simplify=FALSE)
```


# References

