---
title: "Regional topographic heterogeneity"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE)
options(width = 300)
library(knitr)
library(raster)
library(rgeos)
library(tidyverse)
library(dplyr)
library(RColorBrewer)
library(leaflet)
library(htmltools)
library(furrr)
future::plan(multiprocess)
library(ggplot2)
library(GGally)
library(viridis)
library(hrbrthemes)
```

# Data pre-processing and fonctions

The buffer zones used in this document have a radius of 5km, 10km, 20km, 50km, 75km and 100km. 

```{r BufferSize}
buffer.radius <- list(5000,10000,20000,50000,75000,100000) # in meters
names(buffer.radius) <- c(str_sub(buffer.radius[1:5],1,-4),"100")
```

Functions used:

```{r FunctionsUsed}
# Extracting values in the buffer zones from the raster (tiff of each climtic variable)
source("../../scripts/Functions/ExtractValuesFromBuffers.R")

# Plotting the regional climatic heterogeneity by provenances (grouped by gene pools)
source("../../scripts/Functions/PlotRegHeteroByprov.R")

# Plotting the regional climatic heterogeneity by radius of the buffer zones
source("../../scripts/Functions/PlotRegHeteroByradius.R")
```

Create a dataframe with the population structure and provenance coordinates:

```{r DataframeCoordProvPopStructure, message =F}
# Provenance coordinates
data <- readRDS(file="../../data/AllDataPhenoClimSoil.RDS")
df <- unique(data[,c("prov","longitude_prov","latitude_prov")])
df <- df[!(df$prov=="ROD"),]
colnames(df) <- c("prov","longitude","latitude")

# Adding proportion of belonging to each gene pool (population structure)
gp <- data[!(is.na(data$Q1)),]
gp <- gp[!(is.na(gp$height)),]
gp <- droplevels(gp) %>%
  group_by(prov) %>% 
  summarise_at(vars(paste0(rep("Q",6),1:6)), mean)  # %>% mutate_at(2:7,funs(round(., 3))) 
for (i in 1:length(gp$prov)){
  gp[i,"mainGP"] <- names(gp[i,2:7])[which.max(apply(gp[i,2:7],MARGIN=2,max))]
}
colnames(gp) <- c("prov","gpNA","gpC","gpCS","gpFA","gpIA","gpSES","mainGP")

df <- left_join(df,gp)
df
```

Creating the buffer zones:

```{r BufferAroundProvenances,eval=F,echo=F}
xyprov <- SpatialPoints(df[,c("longitude","latitude")], proj4string=CRS("+proj=longlat +datum=WGS84"))
buff5 <- buffer(xyprov,width=buffer.radius[[1]],dissolve=F) # 5km
buff10 <- buffer(xyprov,width=buffer.radius[[2]],dissolve=F) # 10km
buff20 <- buffer(xyprov,width=buffer.radius[[3]],dissolve=F) # 20km
buff50 <- buffer(xyprov,width=buffer.radius[[4]],dissolve=F) # 50km
buff75 <- buffer(xyprov,width=buffer.radius[[5]],dissolve=F) # 75km
buff100 <- buffer(xyprov,width=buffer.radius[[6]],dissolve=F) # 100km
```


# Elevation (DEM)

```{r Elevation,fig.height=7,fig.width=14}
clim.var.name="Elevation"
clim.var.raster <- raster("../../data/Topography/SrtmWGS84/MosaicTif/srtm_mosaic.tif")

# df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
# saveRDS(df2,file="../../data/DfPrePlots/dfSRTM.rds")
df2 <- readRDS(file="../../data/DfPrePlots/dfSRTM.rds")


PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Topographic wetness index TWI

NA for the provenance COM when te buffer zones is 5km or 10km. 


```{r TWI,fig.height=7,fig.width=14}
clim.var.name="TWI"
clim.var.raster <- raster("../../data/Topography/TWI/TWI_UTM31N.tif")

df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
saveRDS(df2,file="../../data/DfPrePlots/dfTWI.rds")
df2 <- readRDS(file="../../data/DfPrePlots/dfTWI.rds")

PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```


# Curvature

<!-- What is curvature? https://www.esri.com/arcgis-blog/products/product/imagery/understanding-curvature-rasters/?rmedium=redirect&rsource=blogs.esri.com/esri/arcgis/2010/10/27/understanding-curvature-rasters -->

```{r Curvature,fig.height=7,fig.width=14}
clim.var.name="Curvature"
clim.var.raster <- raster("../../data/Topography/SlopeAspectCurvature/TifsUTM31N/curvature_mosaic_UTM31N.tif")

# df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
# saveRDS(df2,file="../../data/DfPrePlots/dfCurvature.rds")
df2 <- readRDS(file="../../data/DfPrePlots/dfCurvature.rds")

PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Slope

```{r Slope,fig.height=7,fig.width=14}
clim.var.name="Slope"
clim.var.raster <- raster("../../data/Topography/SlopeAspectCurvature/TifsUTM31N/slope_mosaic_UTM31N.tif")

# df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
# saveRDS(df2,file="../../data/DfPrePlots/dfSlope.rds")
df2 <- readRDS(file="../../data/DfPrePlots/dfSlope.rds")

PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Aspect

```{r Aspect,fig.height=7,fig.width=14}
clim.var.name="Aspect"
clim.var.raster <- raster("../../data/Topography/SlopeAspectCurvature/TifsUTM31N/aspect_mosaic_UTM31N.tif")

# df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
# saveRDS(df2,file="../../data/DfPrePlots/dfAspect.rds")
df2 <- readRDS(file="../../data/DfPrePlots/dfAspect.rds")

PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```