---
title: "Regional climatic heterogeneity"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE)
options(width = 300)
library(knitr)
library(raster)
library(rgeos)
library(tidyverse)
library(dplyr)
library(RColorBrewer)
library(leaflet)
library(htmltools)
library(furrr)
future::plan(multiprocess)
library(ggplot2)
library(GGally)
library(viridis)
library(hrbrthemes)
```

# Data pre-processing and fonctions

The buffer zones used in this document have a radius of 5km, 10km, 20km, 50km, 75km and 100km. 

```{r BufferSize}
buffer.radius <- list(5000,10000,20000,50000,75000,100000) # in meters
names(buffer.radius) <- c(str_sub(buffer.radius[1:5],1,-4),"100")
```

Functions used:


```{r FuntionsUsed}
# Extracting values in the buffer zones from the raster (tiff of each climtic variable)
source("../../scripts/Functions/ExtractValuesFromBuffers.R")

# Plotting the regional climatic heterogeneity by provenances (grouped by gene pools)
source("../../scripts/Functions/PlotRegHeteroByprov.R")

# Mapping with leaflet of the climatic variables and the buffer zones around each provenance
source("../../scripts/Functions/MapClimRegHeteroInteractive.R")

# Plotting the regional climatic heterogeneity by radius of the buffer zones
source("../../scripts/Functions/PlotRegHeteroByradius.R")
```

Create a dataframe with the population structure and provenance coordinates:

```{r DataframeCoordProvPopStructure, message =F}
# Provenance coordinates
data <- readRDS(file="../../data/AllDataPhenoClimSoil.RDS")
df <- unique(data[,c("prov","longitude_prov","latitude_prov")])
df <- df[!(df$prov=="ROD"),]
colnames(df) <- c("prov","longitude","latitude")

# Adding proportion of belonging to each gene pool (population structure)
gp <- data[!(is.na(data$Q1)),]
gp <- gp[!(is.na(gp$height)),]
gp <- droplevels(gp) %>%
  group_by(prov) %>% 
  summarise_at(vars(paste0(rep("Q",6),1:6)), mean)  # %>% mutate_at(2:7,funs(round(., 3))) 
for (i in 1:length(gp$prov)){
  gp[i,"mainGP"] <- names(gp[i,2:7])[which.max(apply(gp[i,2:7],MARGIN=2,max))]
}
colnames(gp) <- c("prov","gpNA","gpC","gpCS","gpFA","gpIA","gpSES","mainGP")

df <- left_join(df,gp)
df
```

Creating the buffer zones:

```{r BufferAroundProvenances}
xyprov <- SpatialPoints(df[,c("longitude","latitude")], proj4string=CRS("+proj=longlat +datum=WGS84"))
buff5 <- buffer(xyprov,width=buffer.radius[[1]],dissolve=F) # 5km
buff10 <- buffer(xyprov,width=buffer.radius[[2]],dissolve=F) # 10km
buff20 <- buffer(xyprov,width=buffer.radius[[3]],dissolve=F) # 20km
buff50 <- buffer(xyprov,width=buffer.radius[[4]],dissolve=F) # 50km
buff75 <- buffer(xyprov,width=buffer.radius[[5]],dissolve=F) # 75km
buff100 <- buffer(xyprov,width=buffer.radius[[6]],dissolve=F) # 100km
```

Labels of the plots:

```{r LabelsProvPopStructure}
labels <- sprintf("<strong> %s </strong><br/>Northern Africa: %s <br/>Corsica: %s <br/>Central Spain: %s <br/>French Atlantic: %s <br/>Iberian Atlantic: %s <br/>South-eastern Spain: %s",
                  df$prov,round(df$gpNA,2), 
                  round(df$gpC,2),
                  round(df$gpCS,2),
                  round(df$gpFA,2),
                  round(df$gpIA,2),
                  round(df$gpSES,2)) %>%
  lapply(htmltools::HTML)
labelsOnlyProv <- sprintf("<strong> %s </strong><br/>",
                  df$prov) %>%
  lapply(htmltools::HTML)
```


# Annual daily mean temperature (bio1)

```{r Bio1Map,fig.height=8,fig.width=8}
clim.var.name="Mean T° (bio1)"
clim.var="bio1"
reverse=TRUE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r Bio1Plots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

- The regional heterogeneity around the Corsican provenances is well captured when the radius is higher than 20km. 

- The Central Spain has both provenances in regions with highly heterogenous temperatures and homogenous temperatures. 

- Maybe the heterogenous temperatures of the South-eastern Spain is better captured when the radius is around 100km? 


# Mean diurnal temperature range (bio2)

```{r Bio2Map,fig.height=8,fig.width=8}
clim.var.name="Diurnal T° range (bio2)"
clim.var="bio2"
reverse=TRUE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r Bio2Plots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Max temperature of the warmest month (bio5)

```{r Bio5Map,fig.height=8,fig.width=8}
clim.var.name="Max T° (bio5)"
clim.var="bio5"
reverse=TRUE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r Bio5Plots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Min temperature of the coldest month (bio6)

```{r Bio6Map,fig.height=8,fig.width=8}
clim.var.name="Min T° (bio6)"
clim.var="bio6"
reverse=TRUE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r Bio6Plots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Annual precipitation (bio12)

```{r Bio12Map,fig.height=8,fig.width=8}
clim.var.name="Annual precipitation (bio12)"
clim.var="bio12"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r Bio12Plots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Precipitation of the wettest month (bio13)

```{r Bio13Map,fig.height=8,fig.width=8}
clim.var.name="Max T° (bio13)"
clim.var="bio13"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r Bio13Plots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Precipitation of the driest month (bio14)

```{r Bio14Map,fig.height=8,fig.width=8}
clim.var.name="Max T° (bio14)"
clim.var="bio14"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r Bio14Plots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```




# Mean P-PET (ppet.mean)

> Precipitation (P) minus potential evapotranspiration (PET)

```{r MeanPPETMap,fig.height=8,fig.width=8}
clim.var.name="Mean P-PET (ppet.mean)"
clim.var="ppet.mean"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r MeanPPETPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```


# Min P-PET (ppet.min)

> Precipitation (P) minus potential evapotranspiration (PET)

```{r MinPPETMap,fig.height=8,fig.width=8}
clim.var.name="Min P-PET (ppet.min)"
clim.var="ppet.min"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r MinPPETPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```


# Max P-PET (ppet.max)

> Precipitation (P) minus potential evapotranspiration (PET)

```{r MaxPPETMap,fig.height=8,fig.width=8}
clim.var.name="Max P-PET (ppet.max)"
clim.var="ppet.max"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r MaxPPETPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```




# Mean PET (pet.mean)

> Potential evapotranspiration (PET)

```{r MeanPETMap,fig.height=8,fig.width=8}
clim.var.name="Mean PET (pet.mean)"
clim.var="pet.mean"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r MeanPETPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```


# Min PET (pet.min)

> Potential evapotranspiration (PET)

```{r MinPETMap,fig.height=8,fig.width=8}
clim.var.name="Min PET (pet.min)"
clim.var="pet.min"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r MinPETPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```


# Max PET (pet.max)

> Potential evapotranspiration (PET)

```{r MaxPETMap,fig.height=8,fig.width=8}
clim.var.name="Max PET (pet.max)"
clim.var="pet.max"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster,reverse)
```

```{r MaxPETPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Winter precipitation (prec.djf)


```{r PrecdjfMap,fig.height=8,fig.width=8}
clim.var.name="Winter precipitation (prec.djf)"
clim.var="prec.djf"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster=clim.var.raster,reverse)
```

```{r PrecdjfPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Spring precipitation (prec.mam)


```{r PrecmamMap,fig.height=8,fig.width=8}
clim.var.name="Spring precipitation (prec.mam)"
clim.var="prec.mam"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster=clim.var.raster,reverse)
```

```{r PrecmamPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Summer precipitation (prec.jja)

```{r PrecjjaMap,fig.height=8,fig.width=8}
clim.var.name="Summer precipitation (prec.jja)"
clim.var="prec.jja"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster=clim.var.raster,reverse)
```

```{r PrecjjaPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Autumn precipitation (prec.son)


```{r PrecsonMap,fig.height=8,fig.width=8}
clim.var.name="Autumn precipitation (prec.son)"
clim.var="prec.son"
reverse=FALSE

clim.var.raster.new <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster=clim.var.raster.new,reverse)
```

```{r PrecsonPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```


# Winter daily mean temperature (tmean.djf)


```{r TmeandjfMap,fig.height=8,fig.width=8}
clim.var.name="Winter T° (tmean.djf)"
clim.var="tmean.djf"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster=clim.var.raster,reverse)
```

```{r TmeandjfPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Spring daily mean temperature (tmean.mam)


```{r TmeanmamMap,fig.height=8,fig.width=8}
clim.var.name="Spring precipitation (tmean.mam)"
clim.var="tmean.mam"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster=clim.var.raster,reverse)
```

```{r TmeanmamPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Summer daily mean temperature (tmean.jja)

```{r TmeanjjaMap,fig.height=8,fig.width=8}
clim.var.name="Summer precipitation (tmean.jja)"
clim.var="tmean.jja"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster=clim.var.raster,reverse)
```

```{r TmeanjjaPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```

# Autumn daily mean temperature (tmean.son)


```{r TmeansonMap,fig.height=8,fig.width=8}
clim.var.name="Autumn precipitation (tmean.son)"
clim.var="tmean.son"
reverse=FALSE

clim.var.raster <- raster(paste0("../../data/climate/Tiff_1901_2009/map_",clim.var,"_1901-2009_with_34_sampling_points.tif"))
MapClimRegHeteroInteractive(clim.var.name,clim.var,clim.var.raster=clim.var.raster,reverse)
```

```{r TmeansonPlots,fig.height=7,fig.width=14}
df2 <- bind_cols(df,future_map_dfr(buffer.radius,ExtractValuesFromBuffers))
PlotRegHeteroByprov(data=df2)
PlotRegHeteroByradius(data=df2)
```




