---
title: "Set of selected environmental variables"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
editor_options: 
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5,fig.height = 4,cache=FALSE)
options(width = 300)
library(knitr)      # CRAN v1.26
library(vegan)      # CRAN v2.5-6    to investigate correlations among predictors
library(psych)      # CRAN v1.8.12   to run RDA and to plot the pairs panels
library(tidyverse)  # CRAN v1.3.0
library(kableExtra) # CRAN v1.1.0
library(raster)     # CRAN v3.3-13 
library(ggbiplot)
```

# Data munging

## Loading phenotypic data

We load the dataset with phenotypic data from a previous paper; script is available here: https://github.com/JulietteArchambeau/H2Pinpin/blob/master/scripts/DataMunging/00_CleaningPhenotypeDataNew.R

In this dataset, clones with no population structure data have already been removed (e.g. ARN1).

<!-- It was updated in November 2019 and before I used the dataset AllDataPhenoClimSoil.RDS with CLONAPIN data of 2018. Some really minor mistakes were corrected between the two versions. -->

```{r LoadPhenoData}
data <- readRDS(file="data/PhenoDataNovember2019.rds")
```



## Merging phenotypic and WorldClim climatic data

We add to the phenotypic data the climatic values in each provenance from WorldClim (from 1970 to 2000).

```{r MergingWorldClimData}
# create a df of the provenance coordinates
xy <- unique(data[,c("prov","longitude_prov","latitude_prov")])

# Path to WorldClim data
path="data/climate/CurrentClimate/WorldClim30sec_1970_2000/"

# List the WorldClim files (the bioclimatic variables)
myFiles <- list.files(path,pattern=".tif")

vars <- str_sub(myFiles,11,-5) %>% str_remove("_") %>% str_c("WC_prov")

for (i in 1:length(myFiles)){
  rast <- raster(paste0(path,myFiles[i]))
  xy[,vars[i]] <- raster::extract(rast,xy[,c("longitude_prov","latitude_prov")])}

data <- merge(data,xy,by=c("prov","longitude_prov","latitude_prov"))

sapply(data,function(x) sum(is.na(x)))
```

Great, no missing data in the climatic variables.


## Merging soil and population structure data

```{r MergeSoilPopStructure}
data2018 <- readRDS(file="data/AllDataPhenoClimSoil.RDS") %>% 
  dplyr::select(clon,prov,contains("top_prov"),"depth_roots_prov",paste0(rep("Q",6),1:6),max.Q) %>% 
  unique()

data <- left_join(data,data2018, by=c("clon","prov"))
data %>% dplyr::select(colnames(data2018)) %>% 
                         sapply(function(x) sum(is.na(x)))
```

Great, no missing data in the columns we merged.


## Merging topography

We include in the dataset the *topographic ruggedness index*.

Topographic data were generated from NASA's Shuttle Radar Topography Mission (SRTM) at 90-m resolution. We used the SAGA v 2.3.1 (Conrad et al. 2015) to calculate the topographic ruggedness index (TRI) which quantifies the terrain heterogeneity, i.e. differences in elevation between adjacent cells (Riley et al. 1999).

```{r MergingTopo}
topo <- readRDS(file="data/Topography/DfTopoVarbyProv.rds")%>% 
  dplyr::select(prov,TRI_WGS84_90m_R) %>% 
  dplyr::rename(TRI=TRI_WGS84_90m_R)

data <- data %>% merge(topo,by="prov")

# checking the missing data
data %>% dplyr::select(colnames(topo)) %>% 
                         sapply(function(x) sum(is.na(x)))
```

Great, no missing data.

## Merging burned area

We include the average of the monthly burned area (in hectares) from June 1995 to December 2014 was extracted from the GFED4 database (Giglio et al., 2013). 

```{r MergingBurnedArea}
BurnedArea <- readRDS(file="data/FireData/DfBurnedAreabyProv.rds") %>% dplyr::select(BurnedArea,prov)

data <- data %>% 
  merge(BurnedArea,by="prov")

# checking the missing data
data %>% dplyr::select(colnames(BurnedArea)) %>% 
                         sapply(function(x) sum(is.na(x)))
```

Great, no missing data.

## Saving the PhD manuscript dataset

As we updated and corrected 

```{r SavingPhDDataset}
# saveRDS(data,"data/PhenoEnvPopStructureDataset.RDS")
```

#  Building the DRYAD dataset

## First paper version

```{r SavingManuscriptDataset}
# We save the dataset for the DRYAD repository
dataDRYAD <- data %>% 
  dplyr::select(-contains("site"), -block,-tree, -obs, -age, -survival, -height,-metap, -region ,-clon_original, -clones522, -block_original, -bb, -dbb, -d13C, -SLA) %>% 
  rename_with(~str_remove(., '_prov')) %>%
  rename_with(~str_remove(., 'WC')) %>%
  unique()

sapply(dataDRYAD,function(x) sum(is.na(x)))

write_csv(dataDRYAD, "data_DRYAD/PopulationPopStructureEnvData.csv", append=FALSE)
```

## Second paper version

We only want the population structure data (Github repository 'GOPredEvalPinpin'). 


```{r WriteCSVpopulationStructure}
data %>% 
  dplyr::select(clon,prov,contains("Q")) %>% 
  distinct() %>% 
  write_csv("../../GOPredEvalPinpin/GOPredEvalPinpin/data_DRYAD/PopulationStructureCorrea2015.csv", append=FALSE)
```

# Selected environmnental variables

Here the set of covariates we are going to use in this paper:

Burned area: `BurnedArea`.

Topographic variable:`TRI`: topographic ruggedness index

Soil variables:

  - `water_top` = sand content in the top layers of the soil
  - `depth_roots` = Depth available to roots

Climatic variables representing episodic events:

  - `bio5`: max temperature of the warmest month (°C)
  - `bio6`: min temperature of the coldest month (°C)
  - `bio12`: annual precipitation (mm)
  - `bio15`: precipitation Seasonality (Coefficient of Variation)


This set of covariates was selected because:

  - in previous analyses based on a subset of the genomic data (Infinium assay data), this is the set (among 6 sets tested) that shows the highest number of candidate SNPs. 
  
  - this set has biological sense. In particular, it includes `bio 6`, the minimum temperature of the coldest month, and studies are accumulating showing that maritime pine patterns of adaptation to climate are associated with minimum temperaturees. 
  
```{r CorrelationsPanels,fig.height=9,fig.width=12}
select.var <- c("bio5WC_prov",
                "bio6WC_prov",
                "bio12WC_prov",
                "bio15WC_prov",
                "water_top_prov",
                "depth_roots_prov",
                "TRI",
                "BurnedArea")

data %>% 
  dplyr::select(all_of(select.var)) %>% 
  unique() %>% 
  dplyr::rename(bio5="bio5WC_prov",
                bio6="bio6WC_prov",
                bio12="bio12WC_prov",
                bio15="bio15WC_prov",
                water_top="water_top_prov",
                depth_roots="depth_roots_prov") %>% 
  pairs.panels(scale=T,hist.col="palegreen1")
```

Other visualizations:

```{r PlotCorr,fig.height=7.3,fig.width=7.3}
source("scripts/Functions/corpmat.R")
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))


PlotCors <- function(data,select.var,nbsize){
  cor <- cor(data[,select.var])

# matrix of the p-value of the correlation
p.mat <- corpmat(cor)

corrplot::corrplot(cor, method="color", col=col(200),  
                   type="upper", order="hclust", 
                   addCoef.col = "black", # Add coefficient of correlation
                   tl.col="black", tl.srt=45, #Text label color and rotation
                   # Combine with significance
                   p.mat = p.mat, sig.level = 0.01, insig = "blank", 
                   # hide correlation coefficient on the principal diagonal
                   diag=FALSE,number.cex=nbsize)

pca <- prcomp(data[,select.var], center = TRUE,scale. = TRUE)

ggbiplot(pca,varname.size =4) +  ylim(-2.5, 2.5) +    xlim(-3, 3) +  
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(size=12))
}

data %>% 
  dplyr::select(all_of(select.var)) %>% 
  unique() %>% 
  dplyr::rename(bio5="bio5WC_prov",
                bio6="bio6WC_prov",
                bio12="bio12WC_prov",
                bio15="bio15WC_prov",
                water_top="water_top_prov",
                depth_roots="depth_roots_prov") %>% 
  PlotCors(select.var = colnames(.),nbsize = 0.6)
```

