---
title: "Fire maps"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
editor_options: 
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5,fig.height = 4,cache=FALSE)
options(width = 300)
library(knitr)
library(tidyverse)
library(kableExtra)
library(raster)
library(caliver)
```

Here a script to download the Fire Weather Index from the following article: https://www.nature.com/articles/sdata201932

The downloading can be done through the package `caliver`:
  
   - paper of the `caliver` package: https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0189419
    
   - tutorial: https://rdrr.io/github/ecmwf/caliver/f/vignettes/caliver_intro-vignette.Rmd
   
   - Data from here: http://globalfiredata.org/pages/data/#burned-area
   
   - Global Fire Emissions Database, Version 4.1 (GFEDv4): https://daac.ornl.gov/cgi-bin/dsviewer.pl?ds_id=1293
   
   - Wildfire Frequency-Area Statistics in Spain: https://www.sciencedirect.com/science/article/pii/S1878029611001599
   
   - Europe: wildfire hazard map: https://www.preventionweb.net/english/professional/maps/v.php?id=3828
   
   - Forests fires map: https://civio.es/en/spain-in-flames/forest-fires-map/

# GFED4 data extraction (a nightmare)
   
Not reproductible at all, work only the first time.... ^^ I had to increase the `Sys.sleep(x)` if I want the function to run again, if I don't, I had the following warning: `Server currently unavailable, please try again later`. This has probably something to do with the repeated number of  connections, the server refuses the access if I try to connect a lot if time in a short amount of time.

So here are the files I have:

First try that worked (extraction from January 1997 to December 2014):

```{bash, eval=F}
cd /home/juliette/Documents/GenomicOffset/GenomicOffsetPinPin/data/FireData/GFED4/Raw
R
source("../../../../scripts/Functions/ExtractBurnedAreasGFED4.R")
BurnedAreas <- get_gfed4(start_date = "1997-01-01",
                         end_date = "2014-12-01",
                         temporal_resolution = "monthly",
                         varname = "BurnedArea")
library("raster")
writeRaster(BurnedAreas,"GFED4_199701_201512.grd", format="raster",overwrite=TRUE)                     
```

Second try that worked, with the whole period of interest, from July 1995 to December 2014.

```{bash, eval=F}
cd /home/juliette/Documents/GenomicOffset/GenomicOffsetPinPin/data/FireData/GFED4/Raw
R
source("../../../../scripts/Functions/ExtractBurnedAreaMonthlyGFED4.R")
rast <- CreateRasterBurnedAreas(start_date = "1995-07-01",end_date = "2014-12-01")
library("raster")
writeRaster(rast,"GFED4_199507_201412.grd", format="raster",overwrite=TRUE)

# saving like in the caliver tutorial:
# https://rdrr.io/github/ecmwf/caliver/f/vignettes/caliver_intro-vignette.Rmd
writeRaster(rast,"GFED4_199507_201412_caliveroptions.grd", format="raster",bandorder='BIL', overwrite=TRUE, progress = 'text') 
```

# Data formating
   
```{r selectArea, eval=F}
Spain <- raster::getData(name = "GADM", country = "Spain", path="data/FireData/GFED4/Raw/",level = 0)
France <- raster::getData(name = "GADM", country = "France", path="data/FireData/GFED4/Raw/", level = 0)
Italy <- raster::getData(name = "GADM", country = "Italy", path="data/FireData/GFED4/Raw/", level = 0)
Morocco <- raster::getData(name = "GADM", country = "Morocco", path="data/FireData/GFED4/Raw/", level = 0)
Algeria <- raster::getData(name = "GADM", country = "Algeria", path="data/FireData/GFED4/Raw/", level = 0)
Portugal <- raster::getData(name = "GADM", country = "Portugal", path="data/FireData/GFED4/Raw/", level = 0)

countries <- rgeos::gUnion(Morocco,Italy) %>% 
  rgeos::gUnion(France) %>% 
  rgeos::gUnion(Spain)  %>% 
  rgeos::gUnion(Algeria) %>% 
  rgeos::gUnion(Portugal)

rast.clim <- raster("data/climate/CurrentClimate/EuMedClim_Tiff_1901_2009/map_bio1_1901-2009_with_34_sampling_points.tif")
b <- as(extent(rast.clim), 'SpatialPolygons')
countries <- rgeos::gIntersection(countries,b)
plot(countries)

shapefile(countries, "data/FireData/GFED4/Raw/countries.shp")
```

# Raster for the genomic analyses 

```{r RasterForGenomicAnalyses, eval=F}
countries <- shapefile("data/FireData/GFED4/Raw/countries.shp")
BurnedAreas <- raster::brick("data/FireData/GFED4/Raw/GFED4_199507_201412.grd")

# Applying the scale factor = 0.01
BurnedAreas <- BurnedAreas * 0.01

# Set extent
raster::extent(BurnedAreas) <- raster::extent(-180, 180, -90, 90)

# Assign CRS (WGS84)
raster::crs(BurnedAreas) <-  "+proj=longlat +datum=WGS84 +no_defs"

# Mask and crop burned areas over Europe
BurnedAreas <- mask_crop_subset(r = BurnedAreas, p = countries)

# Mean over the period of interest
BurnedAreas <- calc(BurnedAreas, mean)

plot(BurnedAreas)

writeRaster(BurnedAreas,"data/FireData/GFED4/PreparedRasters/GFED4_199507_201412_AllCountries.grd")
```


```{r VisualizeRasterGenomic,fig.height=8,fig.width=8}
plot(raster("data/FireData/GFED4/PreparedRasters/GFED4_199507_201412_AllCountries.grd"))
```


## Raster for the NFI models

```{r RasterForNFIModels, eval=F}
Spain <- raster::getData(name = "GADM", country = "Spain", path="data/FireData/GFED4/Raw/",level = 0)
France <- raster::getData(name = "GADM", country = "France", path="data/FireData/GFED4/Raw/", level = 0)
BurnedAreas <- raster::brick("data/FireData/GFED4/Raw/GFED4_199507_201412.grd")

# Applying the scale factor = 0.01
BurnedAreas <- BurnedAreas * 0.01

# Set extent
raster::extent(BurnedAreas) <- raster::extent(-180, 180, -90, 90)

# Assign CRS (WGS84)
raster::crs(BurnedAreas) <-  "+proj=longlat +datum=WGS84 +no_defs"


# If observations layers have no date, assign it!
dataDates <- seq.Date(from = as.Date("1995-07-01"),
                      to = as.Date("2014-12-01"), 
                      by = "month")
names(BurnedAreas) <- dataDates

# Mask and crop burned areas over Spain
BaSp <- mask_crop_subset(r = BurnedAreas, p = Spain)
BaFr <- mask_crop_subset(r = BurnedAreas, p = France)

BaSp <- BaSp[[1:162]] # from July 1995 to December 2008 (range of the third Spanih Inventory)
BaSp <- calc(BaSp, mean)

BaFr <- BaFr[[115:length(names(BaFr))]] # from July 1995 to December 2008 (range of the third Spanih Inventory)
BaFr <- calc(BaFr, mean)

rast <- merge(BaFr,BaSp)

# attribute the extent of the other env. variables (cliamtic, topography, soil)
rast<- crop(rast,extent(-10,14,31,50))

writeRaster(rast,"data/FireData/GFED4/PreparedRasters/GFED4_199507_200712_Spain_200501_201412_France.grd",overwrite=T)
```

```{r VisualizeRasterNFI,fig.height=8,fig.width=8}
plot(raster("data/FireData/GFED4/PreparedRasters/GFED4_199507_200712_Spain_200501_201412_France.grd"))
```


`r knitr::opts_chunk$set(eval = F)`
   

