---
title: "Visualizing spatial differences between current and future climates"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
editor_options: 
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5,fig.height = 4,cache=T)
options(width = 300)
library(knitr)
library(reshape2)
library(ggbiplot)
library(dplyr)
#library(devtools)
#library(xtable)
library(tidyverse)
library(raster)
library(rgeos)
library(kableExtra)
```

```{r FuntionsOptionsUsed,echo=F}
# Gray scale
grayscale_colors <- gray.colors(100,            # number of different color levels
                                start = 0.0,    # how black (0) to go
                                end = 1.0,      # how white (1) to go
                                gamma = 2.2,    # correction between how a digital
                                # camera sees the world and how human eyes see it
                                alpha = NULL)
```


# Load data

## Mapping

For the maps, we need:

  - the maritime pine distribution that we will use as a mask in the maps.
  
  - and hillshade for the background.

```{r LoadMapping}
PinpinDistri  <- shapefile('data/maps/MaskPinpinDistri/PinpinDistriEUforgen_NFIplotsBuffer10km.shp')
background <- raster("data/maps/Hillshade/background.grd")
```


## Environmental data

We load the environmental data:

  - a raster of environmental variables with **current** climate and with **resolution ~ 30.869 km2** (2.5 minutes) (5.556 x 5.556 at the equator), named: `stackCurrClim`.
  
  - a raster of environmental variables with **future** climate under scenario **SSP3-7.0** and with **resolution ~ 30.869 km2** (2.5 minutes), named: `stackFutClimSSP370`.
    
  - a raster of environmental variables with **future** climate under scenario **SSP5-8.5** and with **resolution ~ 30.869 km2** (2.5 minutes), named: `stackFutClimSSP585`.
  

```{r LoadEnvData}
stackCurrClim <- stack("data/StacksEnvVars/Stack_CurrentClimate_Res25minutes.grd")
stackFutClimSSP370 <- stack("data/StacksEnvVars/FutureClimate/2041to2060/SSP370/StackGCMsAvg.grd")
stackFutClimSSP585 <- stack("data/StacksEnvVars/FutureClimate/2041to2060/SSP585/StackGCMsAvg.grd")
```

# Current vs SSP3-7.0

```{r CreatingMapsSSP370}
vars <- c("bio5","bio6","bio12","bio15")

lapply(vars, function(x){
minv <- min(c(min(values(stackCurrClim[[x]]),na.rm=T),min(values(stackFutClimSSP370[[x]]),na.rm=T)))
maxv <- max(c(max(values(stackCurrClim[[x]]),na.rm=T),max(values(stackFutClimSSP370[[x]]),na.rm=T)))

if(x %in% c("bio5","bio6")){
  col=colorRampPalette(c("green","blue","yellow","red","black"))
} else {
  col=colorRampPalette(c("black","red","yellow","blue","green"))
}

pdf(paste0("outputs/DiffEnvCurrentFutureClimate/SSP370/",x,".pdf"),height=12,width=12)
layout(matrix(1:4,2,2))

# current
plot(background,col=grayscale_colors, axes=FALSE,legend=F,main="Current")
plot(stackCurrClim[[x]],zlim=c(minv,maxv) ,add=T,col=col(255))

# future
plot(background,col=grayscale_colors, axes=FALSE,legend=F,main="Future")
plot(stackFutClimSSP370[[x]],add=T,zlim=c(minv,maxv) ,col=col(255))

# absolute difference
diff <- overlay(stackFutClimSSP370[[x]],
                stackCurrClim[[x]],
                fun=function(r1, r2){return(r1-r2)})

plot(background,col=grayscale_colors, axes=FALSE,legend=F,main="Future - Current")
plot(diff,add=T,col=col(255))

# percentage of change. 
diff <- overlay(stackFutClimSSP370[[x]],
                stackCurrClim[[x]],
                fun=function(r1, r2){r1*100/r2-100})

plot(background,col=grayscale_colors, axes=FALSE,legend=F ,
     main="Percentage change btw future and current")
plot(diff,
     add=T,
     col=col(255))
dev.off()


# Map only current climatic conditions
pdf(paste0("outputs/DiffEnvCurrentFutureClimate/SSP370/",x,"_Current.pdf"), 
    width=18, height=18.5)
plot(background,
     col=grayscale_colors,
     axes=FALSE,legend=F)
plot(stackCurrClim[[x]],zlim=c(minv,maxv) ,add=T,col=col(255))
dev.off()

# Map only future climatic conditions
pdf(paste0("outputs/DiffEnvCurrentFutureClimate/SSP370/",x,"_Future.pdf"), 
    width=18, height=18.5)
plot(background,
     col=grayscale_colors,
     axes=FALSE,legend=F)
plot(stackFutClimSSP370[[x]],zlim=c(minv,maxv) ,add=T,col=col(255))
dev.off()


})
```



# Current vs SSP5-8.5

```{r CreatingMaps570}
vars <- c("bio5","bio6","bio12","bio15")

lapply(vars, function(x){
minv <- min(c(min(values(stackCurrClim[[x]]),na.rm=T),min(values(stackFutClimSSP585[[x]]),na.rm=T)))
maxv <- max(c(max(values(stackCurrClim[[x]]),na.rm=T),max(values(stackFutClimSSP585[[x]]),na.rm=T)))

pdf(paste0("outputs/DiffEnvCurrentFutureClimate/SSP585/",x,".pdf"),height=12,width=12)
layout(matrix(1:4,2,2))

# current
plot(background,col=grayscale_colors, axes=FALSE,legend=F,main="Current")
plot(stackCurrClim[[x]],zlim=c(minv,maxv) ,add=T,col=colorRampPalette(c("black","red","yellow","blue","green"))(255))

# future
plot(background,col=grayscale_colors, axes=FALSE,legend=F,main="Future")
plot(stackFutClimSSP585[[x]],add=T,zlim=c(minv,maxv) ,col=colorRampPalette(c("black","red","yellow","blue","green"))(255))

# absolute difference
diff <- overlay(stackFutClimSSP585[[x]],
                stackCurrClim[[x]],
                fun=function(r1, r2){return(r1-r2)})

plot(background,col=grayscale_colors, axes=FALSE,legend=F,main="Future - Current")
plot(diff,add=T,col=colorRampPalette(c("black","red","yellow","blue","green"))(255))

# percentage of change. 
diff <- overlay(stackFutClimSSP585[[x]],
                stackCurrClim[[x]],
                fun=function(r1, r2){r1*100/r2-100})

plot(background,col=grayscale_colors, axes=FALSE,legend=F ,
     main="Percentage change btw future and current")
plot(diff,
     add=T,
     col=colorRampPalette(c("black","red","yellow","blue","green"))(255))
dev.off()
})
```

